---
title: "BeeSearch initial"
author: "Dr. Riley M. Anderson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

What is this analysis about?

### Summary of Results
* 

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(vegan)
library(adespatial)
library(knitr)
library(fossil)

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
dat1 <- read.csv("data/CABS1.csv")

comp.sites <- read.csv("data/comparable.sites.csv")


####################################
# Funtions
####################################

source("functions/TES.R")


```


```{r Data_Wrangling, echo = F, comment = ""}

# general cleaning and conversion:
dat1 <- dat1 %>% 
  mutate(across(c(Station : Sex, Site), as.factor),
         Date = ymd(Label.Date),
         Year = year(Date),
         Month = month(Date),
         # JD --- Day of year (julian day)
         JD = yday(Date),
         # ToY --- Time of Year
         ToY = case_when(JD < 135 ~ "early",
                         JD > 205 ~ "late",
                         TRUE ~ "mid"),
         StationYear = paste(Station, Year, sep = "_")) %>% 
  filter(Station != "") %>% 
  mutate(Sex = case_when(
      Sex == "M" ~ "male",
      Sex == "F" ~ "female",
      TRUE ~ NA_character_ # Handle any unexpected values
    ))


# Sites with comparable sampling:
comp.sites <- comp.sites %>% 
  mutate(StationYear = paste(Station, Year, sep = "_"))

# filter out the sites with incompatible sampling efforts:
comp.data <- semi_join(dat1, comp.sites, by = "StationYear")

comp.data <- comp.data %>% 
  mutate(across(c(Year, StationYear), as.factor))



##################################################
# matrix data

mat1 <- comp.data %>%
  group_by(StationYear, Short.label.name) %>%
  summarise(n = sum(Count)) %>%
  ungroup() %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("StationYear")

mat1meta <- comp.data %>% 
  select(StationYear, Year) %>% 
  distinct()


## matrices by year:

get_matrix_sites_within_years <- function(input_year){
  matrix.site.years <- comp.data %>% 
  filter(Year == input_year) %>% 
  group_by(Station, Short.label.name) %>% 
  summarise(n = sum(Count)) %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n) %>% 
  replace(is.na(.), 0) %>% 
  column_to_rownames("Station")
  
  return(matrix.site.years)
}

matrix.site.years.list <- lapply(levels(comp.data$Year),
                                 get_matrix_sites_within_years)



## matrices by year and time of season:
mat2017 <- comp.data %>% 
  filter(Year == "2017") %>% 
  group_by(Station, Short.label.name, ToY) %>%
  summarise(n = sum(Count)) %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n) %>% 
  replace(is.na(.), 0) %>% 
  mutate(StationTime = paste(Station, ToY, sep = "_")) %>% 
  column_to_rownames("StationTime") %>% 
  select(-Station, -ToY)

meta2017 <- comp.data %>% 
  filter(Year == "2017") %>% 
  select(Station, ToY) %>% 
  distinct() %>% 
  mutate(StationTime = paste(Station, ToY, sep = "_")) %>% 
  column_to_rownames("StationTime")


## matrix by time of season:
mat.seasons <- comp.data %>% 
  filter(Year != "2014" & Year != "2020") %>% 
  group_by(Station, Short.label.name, ToY, Year) %>%
  summarise(n = sum(Count)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n) %>% 
  replace(is.na(.), 0) %>% 
  mutate(StationTimeYear = paste(Station, ToY, Year, sep = "_")) %>% 
  column_to_rownames("StationTimeYear") %>% 
  select(-Station, -ToY, -Year)

meta.seasons <- comp.data %>% 
  filter(Year != "2014" & Year != "2020") %>% 
  select(Station, ToY, Year) %>% 
  distinct() %>% 
  mutate(StationTimeYear = paste(Station, ToY, Year, sep = "_")) %>% 
  column_to_rownames("StationTimeYear")


```

# How many species do we have at each location? How much variability is there in richness?

```{r Q1, echo = F}

rich.year <- comp.data %>% 
  select(Year, Short.label.name) %>% 
  group_by(Year) %>% 
  distinct() %>% 
  tally()

rich.site <- comp.data %>% 
  select(Station, Short.label.name) %>% 
  group_by(Station) %>% 
  distinct() %>% 
  tally()

rich.year.site <- comp.data %>% 
  select(StationYear, Short.label.name) %>% 
  group_by(StationYear) %>% 
  distinct() %>% 
  tally()

rich.site.year.sep <- comp.data %>% 
  select(Station, Year, Short.label.name) %>% 
  group_by(Station, Year) %>% 
  distinct() %>% 
  tally()

rich.site.year.sep %>% 
  filter(Year != "2014" & Year != "2020") %>% 
  ggplot(aes(x = Year, y = n, color = Station, group = Station)) +
  geom_point() +
  geom_line() +
  labs(y = "Richness") +
  theme_classic()

```
This is purely species counts, no kind of offset for sampling effort (other than excluding 2014 and 2020).



# Sex ratio in trap vs. netting collection method:
```{r sex_ratios_function, echo = F}

sex_ratio_function <- function(data, group){
  data  %>%
    group_by(across(all_of(group))) %>%
    tally(name = "count") %>%
    pivot_wider(names_from = Sex,
                values_from = count,
                values_fill = list(count = 0)) %>%
    mutate(sex_ratio = male/female)
}

```
* Ratios > 1 represent male bias
* Ratios < 1 represent female bias


* Overall:
```{r overall_sex_ratios, echo = F}
kable(sex_ratio_function(dat1, c("Sex", "Collection.Method")),
      format = "markdown", digits = 2)
```

* By site:
```{r sex_ratios_by_site, echo = F}
kable(sex_ratio_function(dat1, c("Sex", "Collection.Method", "Site")),
      format = "markdown", digits = 2)
```

* By season:
```{r sex_ratios_by_season, echo = F}
kable(sex_ratio_function(dat1, c("Sex", "Collection.Method", "ToY")),
      format = "markdown", digits = 2)
```

* By season and site:
```{r sex_ratios_by_season_and_site, echo = F}
kable(sex_ratio_function(dat1, c("Sex", "Collection.Method",
                                 "ToY", "Site")),
      format = "markdown", digits = 2)
```

# Chao indices
```{r chao_indices, echo = F, cache = F}

# Sites by season
chao_site_season <- dat1 %>% 
  unite("SiteSeason", c(Site, ToY), sep = "_") %>% 
  group_by(SiteSeason, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = SiteSeason,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Short.label.name")

chao_site_season <- data.frame(
  chao1 = sapply(seq_along(chao_site_season), function(i){
    chao1(chao_site_season[[i]])}),
  chao2 = sapply(seq_along(chao_site_season), function(i){
    chao2(chao_site_season[[i]])}),
  SiteSeason = colnames(chao_site_season)) %>% 
  separate_wider_delim(SiteSeason, delim = "_",
                       names = c("Site", "Season"))

# Sites
chao_site <- dat1 %>% 
  group_by(Site, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = Site,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Short.label.name")

chao_site <- data.frame(
  chao1 = sapply(seq_along(chao_site), function(i){
    chao1(chao_site[[i]])}),
  chao2 = sapply(seq_along(chao_site), function(i){
    chao2(chao_site[[i]])}),
  Site = colnames(chao_site))

# Stations

chao_stations <- dat1 %>% 
  group_by(Station, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = Station,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Short.label.name")

chao_stations <- data.frame(
  chao1 = sapply(seq_along(chao_stations), function(i){
    chao1(chao_stations[[i]])}),
  chao2 = sapply(seq_along(chao_stations), function(i){
    chao2(chao_stations[[i]])}),
  Station = colnames(chao_stations))


# Station by year

chao_station_year <- dat1 %>% 
  group_by(StationYear, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = StationYear,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Short.label.name")

chao_station_year <- data.frame(
  chao1 = sapply(seq_along(chao_station_year), function(i){
    chao1(chao_station_year[[i]])}),
  chao2 = sapply(seq_along(chao_station_year), function(i){
    chao2(chao_station_year[[i]])}),
  StationYear = colnames(chao_station_year)) %>% 
  separate_wider_delim(StationYear, delim = "_",
                       names = c("Station", "Year"))
```

* Chao indices by site
```{r chao_site, echo = F}
kable(chao_site, format = "markdown", digits = 2)
```

* Chao indices by site and season
```{r chao_site_season, echo = F, cache = F}
kable(chao_site_season, format = "markdown", digits = 2)
```

* Chao indices by station
```{r chao_station, echo = F}
kable(chao_stations, format = "markdown", digits = 2)
```

* Chao indices by station and year
```{r chao_station_year, echo = F}
kable(chao_station_year, format = "markdown", digits = 2)
```

# Does community composition change over time?

```{r composition_over_time, echo = F}

adonis2(avgdist(mat1, sample = 5) ~ mat1meta$Year)

permutest(betadisper(avgdist(mat1, sample = 5), mat1meta$Year))

```
No, overall, communities in the POS sites have similar composition across years.

# How does community composition change across seasons?
```{r composition_across_seasons, echo = F, cache=F}

adonis2(avgdist(mat.seasons, sample = 5) ~ meta.seasons$ToY * 
          meta.seasons$Year)

permutest(
  betadisper(
    avgdist(mat.seasons, sample = 5),
    meta.seasons$Year
  )
)


nmds.seasons <- metaMDS(mat.seasons, distance = "bray", k = 2,
                        autotransform = F, try = 250)

stressplot(nmds.seasons)


# site scores
site_scores_seasons <- as.data.frame(scores(nmds.seasons)$sites)

site_scores_seasons <- merge(site_scores_seasons,
                        select(meta.seasons, c(Station, ToY, Year)),
                        by.x = "row.names", by.y = "row.names")

# species scores
spp_scores_seasons <- as.data.frame(scores(nmds.seasons)$species)

spp_scores_seasons <- cbind(spp_scores_seasons,
                       Species = row.names(spp_scores_seasons))
```

```{r nmds_fig_seasons, echo = F, cache=F}

site_scores_seasons <- site_scores_seasons %>%  
  mutate(ToY = factor(ToY)) %>% 
  as.data.frame()


spp_scores_seasons <- as.data.frame(spp_scores_seasons)

centroid_seasons <- site_scores_seasons %>% 
  group_by(ToY) %>% 
  summarise(NMDS1  = mean(NMDS1), NMDS2 = mean(NMDS2))

site_scores_seasons %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = ToY)) +
  geom_point(aes(NMDS1, NMDS2, color = ToY),
             size = 2) +
  stat_ellipse() +
  geom_point(data = centroid_seasons, size = 5, shape = 21,
             color = 'black', aes(fill = ToY), show.legend = F) +
  labs(color = "Time of Season") +
  ggrepel::geom_text_repel(data = spp_scores_seasons,
                           aes(x = NMDS1, y = NMDS2, label = Species),
                           size = 1.5, inherit.aes = FALSE,
                           max.overlaps = 20) +
  theme_cowplot()


```

```{r nmds_fig_years, echo = F, cache=F}

centroid_years <- site_scores_seasons %>% 
  group_by(Year) %>% 
  summarise(NMDS1  = mean(NMDS1), NMDS2 = mean(NMDS2))

site_scores_seasons %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = Year)) +
  geom_point(aes(NMDS1, NMDS2, color = Year),
             size = 2) +
  stat_ellipse() +
  geom_point(data = centroid_years, size = 5, shape = 21,
             color = 'black', aes(fill = Year), show.legend = F) +
  labs(color = "Year") +
  ggrepel::geom_text_repel(data = spp_scores_seasons,
                           aes(x = NMDS1, y = NMDS2, label = Species),
                           size = 1.5, inherit.aes = FALSE,
                           max.overlaps = 20) +
  theme_cowplot()

```


# What bee groups drive changes in abundance from early to mid to late year?


NMDS 2017

```{r NMDS_2017_stress, include = F}

nmds2017 <- metaMDS(mat2017, distance = "bray",
                      k = 2, try = 250, autotransform = F)


nmds2017$dist
stressplot(nmds2017)

```

```{r nmds2017_meta_merge, echo = F}

nmds2017dist <- mat2017 %>% 
  avgdist(sample = 5)


# site scores
site_scores2017 <- as.data.frame(scores(nmds2017)$sites)

site_scores2017 <- merge(site_scores2017,
                        select(meta2017, c(Station, ToY)),
                        by.x = "row.names", by.y = "row.names")

# species scores
spp_scores2017 <- as.data.frame(scores(nmds2017)$species)

spp_scores2017 <- cbind(spp_scores2017,
                       Species = row.names(spp_scores2017))


adonis2(nmds2017dist ~ meta2017$ToY)
permutest(betadisper(nmds2017dist, meta2017$ToY))


```

Within the Port of Seattle sites in 2017, species composition changes significantly throughout the season with distinct groups of species in early, mid, and late season sampling.

```{r nmds2017_plot, echo = F}
site_scores2017 <- site_scores2017 %>%  
  mutate(ToY = factor(ToY)) %>% 
  as.data.frame()


spp_scores2017 <- as.data.frame(spp_scores2017)

centroid2017 <- site_scores2017 %>% 
  group_by(ToY) %>% 
  summarise(NMDS1  = mean(NMDS1), NMDS2 = mean(NMDS2))

site_scores2017 %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = ToY)) +
  geom_point(aes(NMDS1, NMDS2, color = ToY),
             size = 2) +
  stat_ellipse() +
  geom_point(data = centroid2017, size = 5, shape = 21,
             color = 'black', aes(fill = ToY), show.legend = F) +
  labs(color = "Time of Season") +
  ggrepel::geom_text_repel(data = spp_scores2017,
                           aes(x = NMDS1, y = NMDS2, label = Species),
                           size = 1.5, inherit.aes = FALSE,
                           max.overlaps = 20) +
  theme_cowplot()


```
**Non-metric multidimensional scaling of bee species in Port of Seattle in 2017.** Points are sub-sites within the port of Seattle in 2017. They are separated by time of season with the 8 sub sites at the early season in red, the same 8 sub sites at the middle season in blue, and the same 8 sub sites at the end of the season in green. The NMDS space represents total bee species composition and the labels are specific bee species and their position in NMDS space. Large points are the centroids (means) of the points in NMDS space. Ellipses are 95% confidence intervals around the centroids. 

# Can we do analysis with and without major species (H. tripartitus; A. texanus; M. microsticta/us)

What analysis?


# What is the turnover of species within sub-sites within years?

```{r species_turnover_site_years,echo = F}

bdiv.list <- lapply(seq_along(matrix.site.years.list), function(i){
  beta.div.comp(matrix.site.years.list[[i]],
                coef = "J", quant = T)
})

extract_and_combine <- function(list_of_lists, part_name) {
  # Extract the specified part from each sublist and combine into a single data frame
  combined_df <- map2_df(list_of_lists, seq_along(list_of_lists), ~ {
    part_df <- .x[[part_name]]
    part_df <- as.data.frame(part_df)
    part_df <- part_df %>%
      mutate(ListID = .y)  # Add a unique identifier column
    
    return(part_df)
  })
  
  return(combined_df)
}

site.year.part <- extract_and_combine(bdiv.list, "part")

site.year.part <- site.year.part %>% 
  mutate(year = factor(ListID + 2013)) %>% 
  rownames_to_column(var = "metric") %>% 
  mutate(metric = str_remove(metric, "\\.\\.\\..*$")) %>% 
  select(-ListID) %>% 
  pivot_wider(names_from = metric,
              values_from = part_df)

```

```{r POS_beta_div_across_sites_within_years, echo = F}

inset <- site.year.part %>% 
  pivot_longer(cols = c(`Repl/BDtotal`, `RichDif/BDtotal`),
               names_to = "type.contr",
               values_to = "contribution") %>% 
  ggplot(aes(x = year, y = contribution, fill = type.contr)) +
  geom_col() +
  theme_cowplot() +
  scale_fill_discrete(labels = c("species\nturnover",
                                 "richness\ndifferences"),
                      name = "") +
  labs(x = "", y = "Contribution to\nBeta Diversity") +
  theme(axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8, angle = 45),
        axis.title = element_text(size = 10))

plot.y <- site.year.part %>% 
  mutate(year = as.numeric(year) + 2013) %>% 
  ggplot(aes(x = year, y = BDtotal)) +
  geom_line() +
  theme_cowplot() +
  ylim(.24, .40) +
  labs(x = "Year", y = "Beta diversity")

ggdraw(plot.y) +
  draw_plot(inset, .25, .49, .7, .45)

```


# Is there greater overall diversity at SCL vs POS vs PBF?
```{r diversity_among_sites, echo = F}
# Site level matrix
site_matrix <- dat1 %>% 
  group_by(Short.label.name, Site) %>% 
  tally(n = "count") %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Site")

h <- diversity(site_matrix) %>% 
  enframe(name = "Site", value = "Shannon")

simp <- diversity(site_matrix, index = "simpson") %>% 
  enframe(name = "Site", value = "Simpson") %>% 
  select(-Site)

invsimp <- diversity(site_matrix, index = "inv") %>% 
  enframe(name = "Site", value = "InvSimpson") %>% 
  select(-Site)

unbias.simp <- simpson.unb(site_matrix) %>% 
  enframe(name = "Site", value = "UnbiasedSimpson") %>% 
  select(-Site)

f.alpha <- fisher.alpha(site_matrix) %>% 
  enframe(name = "Site", value = "FisherAlpha") %>% 
  select(-Site)



kable(bind_cols(h, simp, invsimp, unbias.simp, f.alpha),
      format = "markdown", digits = 2)
```
Diversity is similar across all 3 sites.



## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


