---
title: "BeeSearch initial"
author: "Dr. Riley M. Anderson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

What is this analysis about?

### Summary of Results
* 

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(vegan)

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
dat1 <- read.csv("data/CABS1.csv")

comp.sites <- read.csv("data/comparable.sites.csv")


####################################
# Funtions
####################################

source("functions/TES.R")


```


```{r Data_Wrangling, echo = F, comment = ""}

# general cleaning and conversion:
dat1 <- dat1 %>% 
  mutate(across(c(Station : Sex, Site), as.factor),
         Date = ymd(Label.Date),
         Year = year(Date),
         Month = month(Date),
         # DoY --- Day of year
         DoY = yday(Date),
         # ToY --- Time of Year
         ToY = case_when(DoY < 135 ~ "early",
                         DoY > 205 ~ "late",
                         TRUE ~ "mid"),
         StationYear = paste(Station, Year, sep = "_"))

# Sites with comparable sampling:
comp.sites <- comp.sites %>% 
  mutate(StationYear = paste(Station, Year, sep = "_"))

# filter the site with incompatible sampling efforts:
comp.data <- semi_join(dat1, comp.sites, by = "StationYear")

comp.data <- comp.data %>% 
  mutate(across(c(Year, StationYear), as.factor))



##################################################
# matrix data

mat1 <- comp.data %>%
  group_by(StationYear, Short.label.name) %>%
  summarise(n = sum(Count)) %>%
  pivot_wider(names_from = Short.label.name,
              values_from = n) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("StationYear")

# 
# mat.meta <- comp.data %>% 
#   select(StationYear, Year, Station, DoY,
#          Month, ToY, Site) %>% 
#   distinct() %>% 
#   column_to_rownames('StationYear')

## matices by year:

mat2017 <- comp.data %>% 
  filter(Year == "2017") %>% 
  group_by(Station, Short.label.name, ToY) %>%
  summarise(n = sum(Count)) %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n) %>% 
  replace(is.na(.), 0) %>% 
  mutate(StationTime = paste(Station, ToY, sep = "_")) %>% 
  column_to_rownames("StationTime") %>% 
  select(-Station, -ToY)

meta2017 <- comp.data %>% 
  filter(Year == "2017") %>% 
  select(Station, ToY) %>% 
  distinct() %>% 
  mutate(StationTime = paste(Station, ToY, sep = "_")) %>% 
  column_to_rownames("StationTime")

```

# How many species do we have at each location? How much variability is there in richness?

```{r Q1, echo = F}

rich.year <- comp.data %>% 
  select(Year, Short.label.name) %>% 
  group_by(Year) %>% 
  distinct() %>% 
  tally()

rich.site <- comp.data %>% 
  select(Station, Short.label.name) %>% 
  group_by(Station) %>% 
  distinct() %>% 
  tally()

rich.year.site <- comp.data %>% 
  select(StationYear, Short.label.name) %>% 
  group_by(StationYear) %>% 
  distinct() %>% 
  tally()

rich.site.year.sep <- comp.data %>% 
  select(Station, Year, Short.label.name) %>% 
  group_by(Station, Year) %>% 
  distinct() %>% 
  tally()

rich.site.year.sep %>% 
  ggplot(aes(x = Year, y = n, color = Station, group = Station)) +
  geom_point() +
  geom_line() +
  labs(y = "Richness") +
  theme_classic()

```

## Rarefaction curves:

What scale do we want?
```{r Rarefaction, echo = F}

tes1 <- TES(t(mat1[,1]))

plot.TES(tes1)



tes2 <- TES(t(mat1[,2]))

plot.TES(tes2)


tes4 <- TES(t(mat1[,4]))

plot.TES(tes4)

tes5 <- TES(t(mat1[,5]))

plot.TES(tes5)

```



# What bee groups drive changes in abundance from early to mid to late year?

```{r NMDS, echo = F}
# 
# bee.nmds2d <- metaMDS(mat1, distance = "bray",
#                       k = 2, try = 250, autotransform = F)
# 
# bee.nmds3d <- metaMDS(mat1, distance = "bray",
#                       k = 3, try = 250, autotransform = F)
# 
# bee.nmds2d$stress
# stressplot(bee.nmds2d)

```

NMDS 2017

```{r NMDS_2017_stress, include = F}

nmds2017 <- metaMDS(mat2017, distance = "bray",
                      k = 2, try = 250, autotransform = F)


nmds2017$stress
stressplot(nmds2017)

```

```{r nmds2017_meta_merge, echo = F}

# site scores
site_scores2017 <- as.data.frame(scores(nmds2017)$sites)

site_scores2017 <- merge(site_scores2017,
                        select(meta2017, c(Station, ToY)),
                        by.x = "row.names", by.y = "row.names")

# species scores
spp_scores2017 <- as.data.frame(scores(nmds2017)$species)

spp_scores2017 <- cbind(spp_scores2017,
                       Species = row.names(spp_scores2017))


```

```{r nmds2017_plot, echo = F}


site_scores2017 %>%  
  ggplot(aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(NMDS1, NMDS2, color = ToY),
             size = 4) +
  labs(color = "Time of Season") +
  ggrepel::geom_text_repel(data = spp_scores2017,
                           aes(x = NMDS1, y = NMDS2, label = Species),
                           size = 2.0) +
  theme_cowplot()

```


# Can we do analysis with and without major species (H. tripartitus; A. texanus; M. microsticta/us)

## - Is there anything we can do to figure out why these species are abundant in certain year?


# What is the turnover of species within sub-sites within years?

# What is the turnover of species within major sites across years?

# How similar are populations within sites based on distance (spatial autocorrelation), and can we assess this for particular bee groups

## Interesting groups: Osmia (lots of species, not abundant), Halictus

# Can we assess why groups like Halictus have so much variability in abundance, and whether this has anything to do with the sex of the individuals collected?

# Is there greater overall diversity at SCL vs POS vs PBF?

# Distance effects – can we estimate flight

# Can we compare to Eli’s study?

# What species are persistent?


## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


