---
title: "BeeSearch initial"
author: "Dr. Riley M. Anderson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = F,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

What is this analysis about?

### Summary of Results
* 

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(vegan)
library(adespatial)
library(knitr)
library(fossil)
library(geosphere)
library(randomForest)
library(caret)


# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
dat1 <- read.csv("data/CABS1.csv")

comp.sites <- read.csv("data/comparable.sites.csv")

site.locs <- read.csv("data/site.locations.csv")
####################################
# Funtions
####################################

source("functions/TES.R")


```


```{r Data_Wrangling, echo = F, comment = ""}

# general cleaning and conversion:
dat1 <- dat1 %>% 
  mutate(across(c(Station : Sex, Site), as.factor), # convert to factors
         Date = ymd(Label.Date), # Convert to date
         Year = year(Date), # Create year column
         Month = month(Date), # create month column
         # JD --- Day of year (julian day)
         JD = yday(Date),
         # ToY --- Time of Year
         ToY = case_when(JD < 135 ~ "early",
                         JD > 205 ~ "late",
                         TRUE ~ "mid"),
         # Create a station/year column for grouping
         StationYear = paste(Station, Year, sep = "_")) %>% 
  filter(Station != "",
         Site != "") %>% # remove empty station & site records
  # convert M/F to male/female to avoid "F = FALSE" misinterpretation
  mutate(Sex = case_when(
      Sex == "M" ~ "male",
      Sex == "F" ~ "female",
      TRUE ~ NA_character_ # Handle any unexpected values
    )) %>% 
  # correct misidentification record
  mutate(Short.label.name = case_when(
    Short.label.name == "Megachile apicalis" ~ "Megachile rotundata",
    TRUE ~ as.character(Short.label.name)
  )) %>% 
  # correct misindentification sex classification
  mutate(Sex = case_when(
    Unique.Specimen.Number == 678 |
      Unique.Specimen.Number == 679 ~ "Male",
    TRUE ~ as.character(Sex)
  ))


# Create data with all the incompatible sites (unequal sampling effort)
incomp.sites <- dat1 %>% 
  select(StationYear) %>% 
  filter(StationYear == "POS1_2018" |
           StationYear == "POS10_2019" |
           StationYear == "POS11_2018" |
           StationYear == "POS13_2017" |
           StationYear == "POS14_2017" |
           StationYear == "POS15_2017" |
           StationYear == "POS16_2017" |
           StationYear == "POS17_2017" |
           StationYear == "POS18_2017" |
           StationYear == "POS19_2017" |
           StationYear == "POS20_2017" |
           StationYear == "POS20_2019" |
           StationYear == "POS21_2017" |
           StationYear == "POS22_2019" |
           StationYear == "POS23_2019" |
           StationYear == "POS1_2014" |
           StationYear == "POS2_2014" |
           StationYear == "POS3_2014") %>% 
  distinct() %>% 
  mutate(StationYear = factor(StationYear))


# filter out the sites with incompatible sampling effort
dat2 <- anti_join(dat1, incomp.sites, by = "StationYear") %>% 
  # unequal sampling in 2020
  filter(Year != 2020) %>% 
  mutate(Year = factor(Year)) %>% 
  droplevels()

# keep the single record of morphospecies "Hylaeus sp. 2023_1"
keep.Hylaeus <- dat2 %>% 
  filter(str_detect(Short.label.name, "Hylaeus sp. 2023"))

# remove any records where species is undetermined:
dat2 <- dat2 %>% 
  filter(!str_detect(Short.label.name, "(?<!m)sp\\.")) %>% 
  bind_rows(keep.Hylaeus)
# ^^^ this data will be used when analyses do not depend on number of species


# Filter out certain morphospecies that could inflate richness
dat.ex.msp <- dat2 %>% 
  filter(!(Sex == "male" & str_detect(Short.label.name,
                                    "Sphecodes msp\\.")),
         !(Sex == "female" & str_detect(Short.label.name,
                                       "Osmia msp\\.")),
         !(Sex == "female" & str_detect(Short.label.name,
                                        "Nomada msp\\.")))
# ^^^ this data will be used for analyses involving species counts


######################################################




##################################################
# matrix data

## matrices by year:

get_matrix_sites_within_years <- function(input_year, data){
  matrix.site.years <- data %>% 
  filter(Year == input_year) %>% 
  group_by(Station, Short.label.name) %>% 
  summarise(n = sum(Count)) %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n,
              values_fill = list(n = 0))  %>% 
  column_to_rownames("Station")
  
  return(matrix.site.years)
}


## matrices by year and time of season:
mat2017 <- dat2 %>% 
  filter(Year == "2017" & Site == "POS" &
           Collection.Method == "T") %>% 
  group_by(Station, Short.label.name, ToY) %>%
  summarise(n = sum(Count)) %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n) %>% 
  replace(is.na(.), 0) %>% 
  mutate(StationTime = paste(Station, ToY, sep = "_")) %>% 
  column_to_rownames("StationTime") %>% 
  select(-Station, -ToY)

meta2017 <- dat2 %>% 
  filter(Year == "2017" & Site == "POS" &
           Collection.Method == "T") %>% 
  select(Station, ToY) %>% 
  distinct() %>% 
  mutate(StationTime = paste(Station, ToY, sep = "_")) %>% 
  column_to_rownames("StationTime")


## matrix by time of season:
mat.seasons <- dat2 %>% 
  filter(Collection.Method == "T" &
           Site == "POS") %>% 
  group_by(Station, Short.label.name, ToY, Year) %>%
  summarise(n = sum(Count)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n) %>% 
  replace(is.na(.), 0) %>% 
  mutate(StationTimeYear = paste(Station, ToY, Year, sep = "_")) %>% 
  column_to_rownames("StationTimeYear") %>% 
  select(-Station, -ToY, -Year)

meta.seasons <- dat2 %>% 
  filter(Site == "POS" &
           Collection.Method == "T") %>% 
  select(Station, ToY, Year) %>% 
  distinct() %>% 
  mutate(StationTimeYear = paste(Station, ToY, Year, sep = "_")) %>% 
  column_to_rownames("StationTimeYear")


```

# How many species do we have at each location? How much variability is there in richness?

```{r Q1, echo = F}

rich.year <- dat.ex.msp %>% 
  select(Year, Short.label.name) %>% 
  group_by(Year) %>% 
  distinct() %>% 
  tally()

rich.site <- dat.ex.msp %>% 
  select(Station, Short.label.name) %>% 
  group_by(Station) %>% 
  distinct() %>% 
  tally()

rich.year.site <- dat.ex.msp %>% 
  select(StationYear, Short.label.name) %>% 
  group_by(StationYear) %>% 
  distinct() %>% 
  tally()

rich.site.year.sep <- dat.ex.msp %>% 
  select(Station, Year, Short.label.name, Site) %>% 
  group_by(Station, Site, Year) %>% 
  distinct() %>% 
  tally()




plots <- rich.site.year.sep %>%
  split(.$Site) %>%
  purrr::map(~ ggplot(.x, aes(x = Year, y = n,
                              color = Station,
                              group = Station)) +
               geom_point() +
               geom_line() +
               labs(y = "Richness", title = unique(.x$Site)) +
               theme_classic() +
               theme(legend.position = "right"))

plot_grid(plotlist = plots, nrow = 3)

```
Raw species counts (excluding incompatible year/stations with altered sampling efforts). Males and females are included. Collection method includes both trap and net.



# Sex ratio in trap vs. netting collection method:
```{r sex_ratios_function, echo = F}

sex_ratio_function <- function(data, group){
  data  %>%
    group_by(across(all_of(group))) %>%
    tally(name = "count") %>%
    pivot_wider(names_from = Sex,
                values_from = count,
                values_fill = list(count = 0)) %>%
    mutate(sex_ratio = male/female)
}

```
* Ratios > 1 represent male bias
* Ratios < 1 represent female bias


* Overall:
```{r overall_sex_ratios, echo = F}
kable(sex_ratio_function(dat2, c("Sex", "Collection.Method")),
      format = "markdown", digits = 2)
```

* By site:
```{r sex_ratios_by_site, echo = F}
kable(sex_ratio_function(dat2, c("Sex", "Collection.Method", "Site")),
      format = "markdown", digits = 2)

sex_ratio_function(dat2, c("Sex", "Collection.Method", "Site")) %>% 
  ggplot(aes(x = Site, y = sex_ratio, fill = Collection.Method)) +
  geom_col(position = position_dodge()) +
  labs(y = "Sex ratio",
       fill = "Collection\nmethod") +
  scale_fill_discrete(labels = c("Net", "Trap")) +
  theme_cowplot() +
  geom_hline(yintercept = 1,
             linetype = "dashed")

```
**Sex ratios of bees by site and collection method.** Points above the dashed line represent male bias, whereas point below represent female bias.


* By season:
```{r sex_ratios_by_season, echo = F}
kable(sex_ratio_function(dat2, c("Sex", "Collection.Method", "ToY")),
      format = "markdown", digits = 2)
```

* By season and site:
```{r sex_ratios_by_season_and_site, echo = F}
kable(sex_ratio_function(dat2, c("Sex", "Collection.Method",
                                 "ToY", "Site")),
      format = "markdown", digits = 2)

sex_ratio_function(dat2, c("Sex",
                           "Collection.Method",
                           "ToY")) %>% 
  ggplot(aes(x = ToY, y = sex_ratio)) +
  geom_point(aes(fill = Collection.Method),
             size = 3,
             shape = 21) +
  scale_x_discrete(limits = c("early", "mid", "late")) +
  geom_jitter(data = filter(sex_ratio_function(dat2, c("Sex",
                                               "Collection.Method",
                                               "ToY", "Site")),
                            sex_ratio < 5),
             aes(x = ToY, y = sex_ratio, color = Collection.Method),
             width = 0.2,
             show.legend = F,
             shape = 21) +
  labs(x = "Time of season",
       y = "Sex ratio",
       fill = "Collection\nmethod") +
  scale_fill_discrete(labels = c("Net", "Trap")) +
  scale_y_continuous(limits = c(0, 5)) +
  theme_cowplot() +
  geom_hline(yintercept = 1,
             linetype = "dashed")

```
**Sex ratios of bees collected by either net or trap across the sampling season.** Points above the dashed line represent male bias, whereas point below represent female bias. Early season (late March - mid May), mid season (mid May to mid July), and late season (mid July - late September) time windows in the sampling effort are shown. Large solid points are sex ratios calculated at the season level. Small open points are sex ratios calculated at the sub-site level within each season. Small points are spread horizontally for visual clarity.


# Species accumulation curves
```{r SAC_prep, echo = F}
### ALL SITES
sac_all <- dat.ex.msp %>% 
  filter(Collection.Method == "T") %>% 
  mutate(StationYear = factor(StationYear)) %>% 
  group_by(StationYear, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("StationYear")

sac_all_list <- specaccum(sac_all, method = "random", permutations = 100,
          conditioned = T, gamma = "jack1")

### POS
sac_pos <- dat.ex.msp %>% 
  filter(Collection.Method == "T",
         Site == "POS") %>% 
  mutate(StationYear = factor(StationYear)) %>% 
  group_by(StationYear, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("StationYear")

sac_pos_list <- specaccum(sac_pos, method = "random", permutations = 100,
                          conditioned = T, gamma = "jack1")

#### SCL
sac_scl <- dat.ex.msp %>% 
  filter(Collection.Method == "T",
         Site == "SCL") %>% 
  mutate(StationYear = factor(StationYear)) %>% 
  group_by(StationYear, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("StationYear")

sac_scl_list <- specaccum(sac_scl, method = "random", permutations = 100,
                          conditioned = T, gamma = "jack1")

### BPF
sac_bpf <- dat.ex.msp %>% 
  filter(Collection.Method == "T",
         Site == "BPF") %>% 
  mutate(StationYear = factor(StationYear)) %>% 
  group_by(StationYear, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("StationYear")

sac_bpf_list <- specaccum(sac_bpf, method = "random", permutations = 100,
                          conditioned = T, gamma = "jack1")

# Results
sac_all_result <- data.frame(site = "All Sites",
                             nsamples = seq(1,58, 1),
                             richness = sac_all_list[4],
                             sd = sac_all_list[5]) %>% 
  mutate(lower = richness - 1.96*sd,
         upper = richness + 1.96*sd)

sac_pos_result <- data.frame(site = "POS",
                             nsamples = seq(1,31, 1),
                             richness = sac_pos_list[4],
                             sd = sac_pos_list[5]) %>% 
  mutate(lower = richness - 1.96*sd,
         upper = richness + 1.96*sd)

sac_scl_result <- data.frame(site = "SCL",
                             nsamples = seq(1,13, 1),
                             richness = sac_scl_list[4],
                             sd = sac_scl_list[5]) %>% 
  mutate(lower = richness - 1.96*sd,
         upper = richness + 1.96*sd)

sac_bpf_result <- data.frame(site = "BPF",
                             nsamples = seq(1,14, 1),
                             richness = sac_bpf_list[4],
                             sd = sac_bpf_list[5]) %>% 
  mutate(lower = richness - 1.96*sd,
         upper = richness + 1.96*sd)


sac_results <- bind_rows(sac_all_result, sac_pos_result,
          sac_scl_result, sac_bpf_result)
```

```{r SAC_figure, echo = F}

sac_results %>% 
  ggplot(aes(x = nsamples, y = richness, fill = site)) +
  geom_line(aes(color = site),linewidth = 1.1, show.legend = F) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = site),
              alpha = 0.4) +
  scale_color_manual(values = c("#000000", "#E69F00", "#56B4E9", "#009E73")) +
  scale_fill_manual(values = c("#999999", "#E69F00", "#56B4E9", "#009E73")) +
  theme_cowplot() +
  labs(x = "No. sampling sites",
       y = "Richness",
       fill = "Site") +
  theme(legend.position = c(0.78,0.4))

```



# Chao indices
```{r chao_indices, echo = F, cache = F}

# Sites by season
chao_site_season <- dat.ex.msp %>% 
  filter(Collection.Method == "T") %>% 
  unite("SiteSeason", c(Site, ToY), sep = "_") %>% 
  group_by(SiteSeason, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = SiteSeason,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Short.label.name")

chao_site_season <- data.frame(
  chao1 = sapply(seq_along(chao_site_season), function(i){
    chao1(chao_site_season[[i]])}),
  SiteSeason = colnames(chao_site_season)) %>% 
  separate_wider_delim(SiteSeason, delim = "_",
                       names = c("Site", "Season"))

# Sites
chao_site <- dat.ex.msp %>% 
  filter(Collection.Method == "T") %>% 
  group_by(Site, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = Site,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Short.label.name")

chao_site <- data.frame(
  chao1 = sapply(seq_along(chao_site), function(i){
    chao1(chao_site[[i]])}),
  Site = colnames(chao_site))

# Stations
chao_stations <- dat.ex.msp %>% 
  filter(Collection.Method == "T") %>% 
  group_by(Station, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = Station,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Short.label.name")

chao_stations <- data.frame(
  chao1 = sapply(seq_along(chao_stations), function(i){
    chao1(chao_stations[[i]])}),
  Station = colnames(chao_stations))


# Station by year
chao_station_year <- dat.ex.msp %>% 
  filter(Collection.Method == "T") %>% 
  group_by(StationYear, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = StationYear,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Short.label.name")

# Station by year (raw values)
station_year_raw <- dat.ex.msp %>% 
  filter(Collection.Method == "T") %>% 
  group_by(StationYear, Short.label.name) %>% 
  tally(name = "count") %>% 
  select(-count) %>% 
  group_by(StationYear) %>% 
  tally(name = "count") %>% 
  separate_wider_delim(StationYear, delim = "_",
                       names = c("Station", "Year")) %>% 
  mutate(Site = substr(Station, 1, 3),
         Year = as.numeric(Year))

# station year (raw means)
station_year_raw_means <- station_year_raw %>% 
  group_by(Site, Year) %>%
  summarise(mean.raw = mean(count))


chao_station_year <- data.frame(
  chao1 = sapply(seq_along(chao_station_year), function(i){
    chao1(chao_station_year[[i]])}),
  StationYear = colnames(chao_station_year)) %>% 
  separate_wider_delim(StationYear, delim = "_",
                       names = c("Station", "Year"))


```
From Anne Chao 1989:

Chao1 minimum species richness is defined non-parametrically as:

$$S_{est} = S_{obs} + f_{1}^{2}/(2f_{2})$$


* Chao indices by site
```{r chao_site, echo = F}
kable(chao_site, format = "markdown", digits = 2)
```

* Chao indices by site and season
```{r chao_site_season, echo = F, cache = F}
kable(chao_site_season, format = "markdown", digits = 2)


chao_site_season %>% 
  mutate(Season = factor(Season)) %>% 
  ggplot(aes(x = Season, y = chao1, color = Site, group = Site)) +
  geom_line(linewidth = 2) +
  scale_x_discrete(limits = c("early", "mid", "late")) +
  theme_cowplot()
  
```

**Overall Chao1 minimum species richness across the sampling season.** Estimated species richness in early (late March - mid May), mid (mid May to mid July), and late (mid July - late September) time windows in the sampling effort. Richness across sites converge in the mid-summer.


* Chao indices by station
```{r chao_station, echo = F}
kable(chao_stations, format = "markdown", digits = 2)
```

* Chao indices by station and year
```{r chao_station_year, echo = F}
kable(chao_station_year, format = "markdown", digits = 2)

chao_station_year %>% 
  mutate(Year = as.numeric(Year),
         Site = substr(Station, 1,3)) %>% 
  group_by(Site, Year) %>% 
  summarise(mean.chao = mean(chao1)) %>% 
  ggplot(aes(x = Year, y = mean.chao,
             color = Site, group = Site)) +
  geom_point(size = 3) +
  geom_line() +
  theme_cowplot() +
  labs(y = "Specie richness") +
  geom_point(data = station_year_raw,
             aes(x = Year, y = count,
                 color = Site),
             shape = 2) +
  geom_line(data = station_year_raw_means,
            aes(x = Year, y = mean.raw, color = Site),
            linetype = 'dashed') +
  ylim(min = 1, max = 85) +
  scale_color_manual(values = c("#E69F00", "#56B4E9", "#009E73")) +
  theme(legend.position = c(0.8, 0.85))




```

**Species richness across sites and sampling years** Points are mean Chao1 estimated species richness, triangles are raw species counts at each substation within each site. Chao richness estimates are lifted by an additive parameter that accounts for rare species likely missed in the sampling. The data exclude some morphospecies (see methods), and all net caught records.



# Does community composition change across seasons, sites, or years?

```{r composition_across_seasons, eval = T, include = F, cache=F}

season.mod <- adonis2(mat.seasons ~ ToY,
        data = meta.seasons, method = "bray")

permutest(betadisper(vegdist(mat.seasons, method = "bray"),
    meta.seasons$ToY))

nmds.seasons <- metaMDS(mat.seasons, distance = "bray", k = 2,
                        autotransform = F, try = 250)

# stressplot(nmds.seasons)
nmds.seasons$stress

# site scores
site_scores_seasons <- as.data.frame(scores(nmds.seasons)$sites)

site_scores_seasons <- merge(site_scores_seasons,
                        select(meta.seasons, c(Station, ToY, Year)),
                        by.x = "row.names", by.y = "row.names")

# species scores
spp_scores_seasons <- as.data.frame(scores(nmds.seasons)$species)

spp_scores_seasons <- cbind(spp_scores_seasons,
                       Species = row.names(spp_scores_seasons))
```


## Across seasons: (I don't think we should include this)
```{r nmds_seasons_prep, eval = T, include = F, cache=F}

site_scores_seasons <- site_scores_seasons %>%  
  mutate(ToY = factor(ToY),
         Year = factor(Year)) %>% 
  as.data.frame()


spp_scores_seasons <- as.data.frame(spp_scores_seasons)

centroid_seasons <- site_scores_seasons %>% 
  group_by(ToY) %>% 
  summarise(NMDS1  = mean(NMDS1), NMDS2 = mean(NMDS2))
```

```{r ndms_fig_seasons, echo = F}
site_scores_seasons %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = ToY)) +
  geom_point(aes(NMDS1, NMDS2, color = ToY),
             size = 2) +
  stat_ellipse() +
  geom_point(data = centroid_seasons, size = 5, shape = 21,
             color = 'black', aes(fill = ToY), show.legend = F) +
  labs(color = "Time of Season") +
  ggrepel::geom_text_repel(data = spp_scores_seasons,
                           aes(x = NMDS1, y = NMDS2, label = Species),
                           size = 1.5, inherit.aes = FALSE,
                           max.overlaps = 20) +
  theme_cowplot()


```
**Community composition across time of season.** Strong overlap of composition across the early, mid, and late season sampling (PERMANOVA: F = `r round(season.mod[1,4], 2)`, `r round(season.mod[1,5], 2)`). Data include all compatible stations across all years.



## Across sites
```{r nmds_all_sites_prep_and_tests, echo = F}

#####################################################
# matrix assembly
site_matrix <- dat2 %>% 
  filter(Collection.Method == "T") %>% 
  group_by(Short.label.name, StationYear) %>% 
  tally(n = "count") %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("StationYear")

site_meta <- dat2 %>% 
  filter(Collection.Method == "T") %>% 
  select(Site, Station, StationYear, Year) %>% 
  distinct()
#####################################################



#############################################################
# Stats
site.year.mod <- adonis2(site_matrix ~ Site * Year,
        data = site_meta, method = "bray")

permutest(betadisper(vegdist(site_matrix, method = "bray"),
                     site_meta$Year))

permutest(betadisper(vegdist(site_matrix, method = "bray"),
                     site_meta$Site))

```


```{r nmds_sites_calcs, eval = T, include = F}
# NMDS scores
nmds_SSY <- metaMDS(site_matrix, distance = "bray",
                    k = 3, try = 250, autotransform = F)
nmds_SSY$stress
# 
# stressplot(nmds_SSY)

site_scores <- data.frame(scores(nmds_SSY)$sites) %>%
  rownames_to_column("StationYear") %>% 
  left_join(site_meta, by = "StationYear") %>%
  unite(SiteYear, c("Site", "Year"), sep = "_", remove = F) %>% 
  mutate(across(c(SiteYear:Year), as.factor)) %>% 
  filter(Site != "") %>% 
  droplevels()

spp_scores <- data.frame(scores(nmds_SSY)$species) %>% 
  rownames_to_column("Species")

centroids_SSY <- site_scores %>% 
  group_by(SiteYear) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))

centroids_site <- site_scores %>% 
  group_by(Site) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))

centroids_year <- site_scores %>% 
  group_by(Year) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))
```

### Site classification by species composition (Random Forest)
```{r random_forest_sites, echo = F}

rf_matrix <- site_matrix %>% 
  rownames_to_column("StationYear") %>% 
  separate_wider_delim(StationYear, delim = "_",
                       cols_remove = F,
                       names = c("Station", "Year")) %>% 
  separate_wider_position(Station,
                          widths = c(Site = 3, 3, StationID = 2),
                          cols_remove = F,
                          too_few = "debug") %>% 
  select(-StationID, -Station_width,
         -Station_remainder, -Station_ok) %>% 
  mutate(across(c(Site:StationYear), as.factor))

rfmod_site <- randomForest(y = rf_matrix$Site,
                           x = select(rf_matrix,
                                      -Year, -Station,
                                      -StationYear, -Site),
                           importance = T, nPerm = 999,
                           proximity = T)


train(y = rf_matrix$Site,
      x = select(rf_matrix,
                 -Year, -Station,
                 -StationYear, -Site),
      method = "rf", na.action = na.omit)

update(rfmod_site, mtry = 68)


imp_vals_top <- data.frame(importance(rfmod_site)) %>% 
  arrange(desc(MeanDecreaseGini)) %>% 
  filter(MeanDecreaseGini > quantile(MeanDecreaseGini, 0.9))

kable(imp_vals_top, format = "markdown", digits = 3)

MDSplot(rfmod_site, rf_matrix$Site)
```

**Random Forest classification of site by species composition.** The model was tuned without pre-processing. Bootstrapped resampling used 25 replicates. Overall model accuracy was 82%. The model clearly delineated the BPF sites with 0.`r rfmod_site$confusion[1,4]` class error. Similarly, POS sites were near perfect with `r round(rfmod_site$confusion[2,4], 3)` class error. However, the SCL sites were less accurately classified (`r round(rfmod_site$confusion[3,4], 3)` class error), with (`r rfmod_site$confusion[3,2]`/13) identified as POS. This lends further support to the NMDS figure below, species composition is very similar across SCL and POS, but BPF sites have different community composition compared to the other two sites. 

The table above shows the species most representative of the community differences used to make the above classifications. These species are the top 10th percentile of ranked variable importance (mean decrease in Gini score).


### NMDS by Site figure (include this)
```{r nmds_all_sites_year, echo = F}

imp_spp <- imp_vals_top %>% 
  rownames_to_column("Species") %>% 
  select(Species, MeanDecreaseGini)

spp_scores <- right_join(spp_scores, imp_spp, by = "Species")

site_scores %>% 
  ggplot(aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(NMDS1, NMDS2, color = Site),
             size = 1) +
  stat_ellipse(aes(color = Site),
               show.legend = F) +
  geom_point(data = centroids_site,
             aes(x = NMDS1, y = NMDS2, fill = Site),
             size = 4,
             shape = 21, show.legend = F) +
  theme_cowplot() +
  ggrepel::geom_text_repel(data = spp_scores,
                           aes(x = NMDS1, y = NMDS2,
                               label = Species,
                               size = MeanDecreaseGini),
                           max.overlaps = 20) +
  scale_size_continuous(
    name = "Mean decrease Gini",
    limits = c(.4, 3),
    breaks = c(0.75, 2, 3),
    range = c(1, 6)) +
  guides(color = guide_legend(override.aes = list(size = 4, shape = 16))) +
  scale_color_manual(values = c("#E69F00", "#56B4E9", "#009E73"))+
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73"))


```
**Variation in community composition across sites.** Bee species are plotted on the first two axes of a three-dimensional non-metric multidimensional scaling ordination of the 58 combinations of station (subsite) and year, across the three sites. Small points are the individual station/year combinations. Large points are the centroids of the three sites. Ellipses are 95% confidence intervals around the site centroids.  Bees species shown are the most representative (top 10th percentile of a random forest analysis) of the compositional differences among sites. Text size of the labels is proportional to variable importance score (mean decrease in Gini score).


## Across years (I don't think we should include this)
```{r nmds_all_year, echo = F}
site_scores %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = Year)) +
  geom_point(aes(NMDS1, NMDS2, color = Year),
             size = 1) +
  geom_point(data = centroids_year,
             aes(x = NMDS1, y = NMDS2, fill = Year),
             shape = 21,
             color = 'black',
             size = 4,
             show.legend = F) +
  stat_ellipse(aes(color = Year)) +
  theme_cowplot()
```
**Community composition across years.** Most years have strong overlap of composition but the *year* term is highly significant (PERMANOVA: F = `r round(site.year.mod[2,4], 2)`, *P* = `r round(site.year.mod[2,5], 3)`). However, the only site in 2014 is SCL and the 2018 and 2019 years are heavily influenced by the BPF data. Essentially, the information in *years* is not sufficiently distinct from the information in *sites* and including *year* in the model is not informative. This is also the likely cause of the violation of homogeneity of multivariate dispersions for the *year* term but not the *site* term.


# POS 2017 Changes over seasons (This is not interesting)

```{r NMDS_2017_stress, include = F}

nmds2017 <- metaMDS(mat2017, distance = "bray",
                      k = 2, try = 250, autotransform = F)

nmds2017$dist
stressplot(nmds2017)

```

```{r nmds2017_meta_merge, echo = F}

nmds2017dist <- mat2017 %>% 
  avgdist(sample = 5)


# site scores
site_scores2017 <- as.data.frame(scores(nmds2017)$sites)

site_scores2017 <- merge(site_scores2017,
                        select(meta2017, c(Station, ToY)),
                        by.x = "row.names", by.y = "row.names")

# species scores
spp_scores2017 <- as.data.frame(scores(nmds2017)$species)

spp_scores2017 <- cbind(spp_scores2017,
                       Species = row.names(spp_scores2017))


adonis2(nmds2017dist ~ meta2017$ToY, method = "bray")
permutest(betadisper(nmds2017dist, meta2017$ToY))


```
Within the Port of Seattle sites in 2017 and only 2017, species composition changes significantly throughout the season with distinct groups of species in early, mid, and late season sampling. **However, this model violates homogeneity of multivariate dispersions.**


```{r nmds2017_plot, echo = F}
site_scores2017 <- site_scores2017 %>%  
  mutate(ToY = factor(ToY)) %>% 
  as.data.frame()


spp_scores2017 <- as.data.frame(spp_scores2017)

centroid2017 <- site_scores2017 %>% 
  group_by(ToY) %>% 
  summarise(NMDS1  = mean(NMDS1), NMDS2 = mean(NMDS2))

site_scores2017 %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = ToY)) +
  geom_point(aes(NMDS1, NMDS2, color = ToY),
             size = 2) +
  stat_ellipse() +
  geom_point(data = centroid2017, size = 5, shape = 21,
             color = 'black', aes(fill = ToY), show.legend = F) +
  labs(color = "Time of Season") +
  ggrepel::geom_text_repel(data = spp_scores2017,
                           aes(x = NMDS1, y = NMDS2, label = Species),
                           size = 1.5, inherit.aes = FALSE,
                           max.overlaps = 20) +
  theme_cowplot()


```
**Non-metric multidimensional scaling of bee species in Port of Seattle in 2017.** Points are sub-sites within the port of Seattle in 2017. They are separated by time of season with the 8 sub sites at the early season in red, the same 8 sub sites at the middle season in blue, and the same 8 sub sites at the end of the season in green. The NMDS space represents total bee species composition and the labels are specific bee species and their position in NMDS space. Large points are the centroids (means) of the points in NMDS space. Ellipses are 95% confidence intervals around the centroids. 



# What is the turnover of species within sub-sites within years?

```{r species_turnover_site_years_POS, echo = F}

POS <- dat2 %>% 
  filter(Site == "POS") %>% 
  droplevels()

pos.matrix.list <- lapply(levels(POS$Year),
                                 data = POS,
                                 get_matrix_sites_within_years)


bdiv.list <- lapply(seq_along(pos.matrix.list), function(i){
  beta.div.comp(pos.matrix.list[[i]],
                coef = "J", quant = T)
})

extract_and_combine <- function(list_of_lists, part_name) {
  # Extract the specified part from each sublist and combine into a single data frame
  combined_df <- map2_df(list_of_lists, seq_along(list_of_lists), ~ {
    part_df <- .x[[part_name]]
    part_df <- as.data.frame(part_df)
    part_df <- part_df %>%
      mutate(ListID = .y)  # Add a unique identifier column
    
    return(part_df)
  })
  
  return(combined_df)
}

site.year.part <- extract_and_combine(bdiv.list, "part")

site.year.part <- site.year.part %>% 
  mutate(year = factor(ListID + 2014)) %>% 
  rownames_to_column(var = "metric") %>% 
  mutate(metric = str_remove(metric, "\\.\\.\\..*$")) %>% 
  select(-ListID) %>% 
  pivot_wider(names_from = metric,
              values_from = part_df)

```

### POS
```{r POS_beta_div_across_sites_within_years, echo = F}

inset <- site.year.part %>% 
  pivot_longer(cols = c(`Repl/BDtotal`, `RichDif/BDtotal`),
               names_to = "type.contr",
               values_to = "contribution") %>% 
  ggplot(aes(x = year, y = contribution, fill = type.contr)) +
  geom_col() +
  theme_cowplot() +
  scale_fill_discrete(labels = c("species\nturnover",
                                 "richness\ndifferences"),
                      name = "") +
  labs(x = "", y = "Contribution to\nBeta Diversity") +
  theme(axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8, angle = 45),
        axis.title = element_text(size = 10))

plot.y <- site.year.part %>% 
  mutate(year = as.numeric(year) + 2014) %>% 
  ggplot(aes(x = year, y = BDtotal)) +
  geom_line() +
  theme_cowplot() +
  ylim(.24, .62) +
  labs(x = "Year", y = "Beta diversity")

ggdraw(plot.y) +
  draw_plot(inset, .25, .49, .7, .45)

```

### SCL
```{r species_turnover_site_years_SCL, echo = F}

SCL <- dat2 %>% 
  filter(Site == "SCL") %>% 
  droplevels()

scl.matrix.list <- lapply(levels(SCL$Year),
                                 data = SCL,
                                 get_matrix_sites_within_years)


bdiv.list.scl <- lapply(seq_along(scl.matrix.list), function(i){
  beta.div.comp(scl.matrix.list[[i]],
                coef = "J", quant = T)
})

site.year.part.scl <- extract_and_combine(bdiv.list.scl, "part")

site.year.part.scl <- site.year.part.scl %>% 
  mutate(year = factor(ListID + 2013)) %>% 
  rownames_to_column(var = "metric") %>% 
  mutate(metric = str_remove(metric, "\\.\\.\\..*$")) %>% 
  select(-ListID) %>% 
  pivot_wider(names_from = metric,
              values_from = part_df)

```

```{r SCL_beta_div_across_sites_within_years, echo = F}

inset.scl <- site.year.part.scl %>% 
  pivot_longer(cols = c(`Repl/BDtotal`, `RichDif/BDtotal`),
               names_to = "type.contr",
               values_to = "contribution") %>% 
  ggplot(aes(x = year, y = contribution, fill = type.contr)) +
  geom_col() +
  theme_cowplot() +
  scale_fill_discrete(labels = c("species\nturnover",
                                 "richness\ndifferences"),
                      name = "") +
  labs(x = "", y = "Contribution to\nBeta Diversity") +
  theme(axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8, angle = 45),
        axis.title = element_text(size = 10))

plot.y.scl <- site.year.part.scl %>% 
  mutate(year = as.numeric(year) + 2013) %>% 
  ggplot(aes(x = year, y = BDtotal)) +
  geom_line() +
  theme_cowplot() +
  ylim(.27, .45) +
  labs(x = "Year", y = "Beta diversity") +
  scale_x_continuous(breaks = c(2014, 2015, 2016))

ggdraw(plot.y.scl) +
  draw_plot(inset.scl, .25, .49, .7, .45)

```


### BPF
```{r species_turnover_site_years_BPF, echo = F}
BPF <- dat2 %>% 
  filter(Site == "BPF") %>% 
  droplevels()

BPF.matrix.list <- lapply(levels(BPF$Year),
                          data = BPF,
                          get_matrix_sites_within_years)

bdiv.list.bpf <- lapply(seq_along(BPF.matrix.list), function(i){
  beta.div.comp(BPF.matrix.list[[i]],
                coef = "J", quant = T)
})

site.year.part.bpf <- extract_and_combine(bdiv.list.bpf, "part")

site.year.part.bpf <- site.year.part.bpf %>% 
  mutate(year = factor(ListID + 2017)) %>% 
  rownames_to_column(var = "metric") %>% 
  mutate(metric = str_remove(metric, "\\.\\.\\..*$")) %>% 
  select(-ListID) %>% 
  pivot_wider(names_from = metric,
              values_from = part_df)

```

```{r BPF_beta_div_across_sites_within_years, echo = F}

inset.bpf <- site.year.part.bpf %>% 
  pivot_longer(cols = c(`Repl/BDtotal`, `RichDif/BDtotal`),
               names_to = "type.contr",
               values_to = "contribution") %>% 
  ggplot(aes(x = year, y = contribution, fill = type.contr)) +
  geom_col() +
  theme_cowplot() +
  scale_fill_discrete(labels = c("species\nturnover",
                                 "richness\ndifferences"),
                      name = "") +
  labs(x = "", y = "Contribution to\nBeta Diversity") +
  theme(axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8, angle = 45),
        axis.title = element_text(size = 10))

plot.y.bpf <- site.year.part.bpf %>% 
  mutate(year = as.numeric(year) + 2017) %>% 
  ggplot(aes(x = year, y = BDtotal)) +
  geom_line() +
  theme_cowplot() +
  ylim(.33, .50) +
  labs(x = "Year", y = "Beta diversity") +
  scale_x_continuous(breaks = c(2018, 2019))

ggdraw(plot.y.bpf) +
  draw_plot(inset.bpf, .25, .49, .7, .45)

```



# Is there greater overall diversity at SCL vs POS vs PBF?
```{r diversity_among_sites, echo = F}
# Site level matrix
site_matrix <- dat1 %>% 
  group_by(Short.label.name, Site) %>% 
  tally(n = "count") %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Site")

h <- diversity(site_matrix) %>% 
  enframe(name = "Site", value = "Shannon")

simp <- diversity(site_matrix, index = "simpson") %>% 
  enframe(name = "Site", value = "Simpson") %>% 
  select(-Site)

invsimp <- diversity(site_matrix, index = "inv") %>% 
  enframe(name = "Site", value = "InvSimpson") %>% 
  select(-Site)

unbias.simp <- simpson.unb(site_matrix) %>% 
  enframe(name = "Site", value = "UnbiasedSimpson") %>% 
  select(-Site)

f.alpha <- fisher.alpha(site_matrix) %>% 
  enframe(name = "Site", value = "FisherAlpha") %>% 
  select(-Site)



kable(bind_cols(h, simp, invsimp, unbias.simp, f.alpha),
      format = "markdown", digits = 2)
```
Diversity is similar across all 3 sites.



## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


