---
title: "BeeSearch initial"
author: "Dr. Riley M. Anderson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

What is this analysis about?

### Summary of Results
* 

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(vegan)
library(adespatial)

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
dat1 <- read.csv("data/CABS1.csv")

comp.sites <- read.csv("data/comparable.sites.csv")


####################################
# Funtions
####################################

source("functions/TES.R")


```


```{r Data_Wrangling, echo = F, comment = ""}

# general cleaning and conversion:
dat1 <- dat1 %>% 
  mutate(across(c(Station : Sex, Site), as.factor),
         Date = ymd(Label.Date),
         Year = year(Date),
         Month = month(Date),
         # JD --- Day of year (julian day)
         JD = yday(Date),
         # ToY --- Time of Year
         ToY = case_when(JD < 135 ~ "early",
                         JD > 205 ~ "late",
                         TRUE ~ "mid"),
         StationYear = paste(Station, Year, sep = "_"))

# Sites with comparable sampling:
comp.sites <- comp.sites %>% 
  mutate(StationYear = paste(Station, Year, sep = "_"))

# filter out the site with incompatible sampling efforts:
comp.data <- semi_join(dat1, comp.sites, by = "StationYear")

comp.data <- comp.data %>% 
  mutate(across(c(Year, StationYear), as.factor))



##################################################
# matrix data

mat1 <- comp.data %>%
  group_by(StationYear, Short.label.name) %>%
  summarise(n = sum(Count)) %>%
  ungroup() %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("StationYear")


## matrices by year:

get_matrix_sites_within_years <- function(input_year){
  matrix.site.years <- comp.data %>% 
  filter(Year == input_year) %>% 
  group_by(Station, Short.label.name) %>% 
  summarise(n = sum(Count)) %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n) %>% 
  replace(is.na(.), 0) %>% 
  column_to_rownames("Station")
  
  return(matrix.site.years)
}

matrix.site.years.list <- lapply(levels(comp.data$Year),
                                 get_matrix_sites_within_years)



## matrices by year and time of season:
mat2017 <- comp.data %>% 
  filter(Year == "2017") %>% 
  group_by(Station, Short.label.name, ToY) %>%
  summarise(n = sum(Count)) %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n) %>% 
  replace(is.na(.), 0) %>% 
  mutate(StationTime = paste(Station, ToY, sep = "_")) %>% 
  column_to_rownames("StationTime") %>% 
  select(-Station, -ToY)

meta2017 <- comp.data %>% 
  filter(Year == "2017") %>% 
  select(Station, ToY) %>% 
  distinct() %>% 
  mutate(StationTime = paste(Station, ToY, sep = "_")) %>% 
  column_to_rownames("StationTime")

```

# How many species do we have at each location? How much variability is there in richness?

```{r Q1, echo = F}

rich.year <- comp.data %>% 
  select(Year, Short.label.name) %>% 
  group_by(Year) %>% 
  distinct() %>% 
  tally()

rich.site <- comp.data %>% 
  select(Station, Short.label.name) %>% 
  group_by(Station) %>% 
  distinct() %>% 
  tally()

rich.year.site <- comp.data %>% 
  select(StationYear, Short.label.name) %>% 
  group_by(StationYear) %>% 
  distinct() %>% 
  tally()

rich.site.year.sep <- comp.data %>% 
  select(Station, Year, Short.label.name) %>% 
  group_by(Station, Year) %>% 
  distinct() %>% 
  tally()

rich.site.year.sep %>% 
  ggplot(aes(x = Year, y = n, color = Station, group = Station)) +
  geom_point() +
  geom_line() +
  labs(y = "Richness") +
  theme_classic()

```

## Rarefaction curves:

What scale do we want?
```{r Rarefaction, echo = F}

tes1 <- TES(t(mat1[,1]))

plot.TES(tes1)



tes2 <- TES(t(mat1[,2]))

plot.TES(tes2)


tes4 <- TES(t(mat1[,4]))

plot.TES(tes4)

tes5 <- TES(t(mat1[,5]))

plot.TES(tes5)

```



# What bee groups drive changes in abundance from early to mid to late year?

```{r NMDS, echo = F}
# 
# bee.nmds2d <- metaMDS(mat1, distance = "bray",
#                       k = 2, try = 250, autotransform = F)
# 
# bee.nmds3d <- metaMDS(mat1, distance = "bray",
#                       k = 3, try = 250, autotransform = F)
# 
# bee.nmds2d$stress
# stressplot(bee.nmds2d)

```

NMDS 2017

```{r NMDS_2017_stress, include = F}

nmds2017 <- metaMDS(mat2017, distance = "bray",
                      k = 2, try = 250, autotransform = F)


nmds2017$dist
stressplot(nmds2017)

```

```{r nmds2017_meta_merge, echo = F}

nmds2017dist <- mat2017 %>% 
  avgdist(sample = 5)


# site scores
site_scores2017 <- as.data.frame(scores(nmds2017)$sites)

site_scores2017 <- merge(site_scores2017,
                        select(meta2017, c(Station, ToY)),
                        by.x = "row.names", by.y = "row.names")

# species scores
spp_scores2017 <- as.data.frame(scores(nmds2017)$species)

spp_scores2017 <- cbind(spp_scores2017,
                       Species = row.names(spp_scores2017))


adonis2(nmds2017dist ~ meta2017$ToY)
permutest(betadisper(nmds2017dist, meta2017$ToY))


```

Within the Port of Seattle sites in 2017, species composition changes significantly throughout the season with distinct groups of species in early, mid, and late season sampling.

```{r nmds2017_plot, echo = F}
site_scores2017 <- site_scores2017 %>%  
  mutate(ToY = factor(ToY)) %>% 
  as.data.frame()


spp_scores2017 <- as.data.frame(spp_scores2017)

centroid2017 <- site_scores2017 %>% 
  group_by(ToY) %>% 
  summarise(NMDS1  = mean(NMDS1), NMDS2 = mean(NMDS2))

site_scores2017 %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = ToY)) +
  geom_point(aes(NMDS1, NMDS2, color = ToY),
             size = 2) +
  stat_ellipse() +
  geom_point(data = centroid2017, size = 5, shape = 21,
             color = 'black', aes(fill = ToY), show.legend = F) +
  labs(color = "Time of Season") +
  ggrepel::geom_text_repel(data = spp_scores2017,
                           aes(x = NMDS1, y = NMDS2, label = Species),
                           size = 1.5, inherit.aes = FALSE,
                           max.overlaps = 20) +
  theme_cowplot()


```
**Non-metric multidimensional scaling of bee species in Port of Seattle in 2017.** Points are sub-sites within the port of Seattle in 2017. They are separated by time of season with the 8 sub sites at the early season in red, the same 8 sub sites at the middle season in blue, and the same 8 sub sites at the end of the season in green. The NMDS space represents total bee species composition and the labels are specific bee species and their position in NMDS space. Large points are the centroids (means) of the points in NMDS space. Ellipses are 95% confidence intervals around the centroids. 

# Can we do analysis with and without major species (H. tripartitus; A. texanus; M. microsticta/us)

## - Is there anything we can do to figure out why these species are abundant in certain year?


# What is the turnover of species within sub-sites within years?

```{r species_turnover_site_years,echo = F}

bdiv.list <- lapply(seq_along(matrix.site.years.list), function(i){
  beta.div.comp(matrix.site.years.list[[i]],
                coef = "J", quant = T)
})

extract_and_combine <- function(list_of_lists, part_name) {
  # Extract the specified part from each sublist and combine into a single data frame
  combined_df <- map2_df(list_of_lists, seq_along(list_of_lists), ~ {
    part_df <- .x[[part_name]]
    part_df <- as.data.frame(part_df)
    part_df <- part_df %>%
      mutate(ListID = .y)  # Add a unique identifier column
    
    return(part_df)
  })
  
  return(combined_df)
}

site.year.part <- extract_and_combine(bdiv.list, "part")

site.year.part <- site.year.part %>% 
  mutate(year = factor(ListID + 2013)) %>% 
  rownames_to_column(var = "metric") %>% 
  mutate(metric = str_remove(metric, "\\.\\.\\..*$")) %>% 
  select(-ListID) %>% 
  pivot_wider(names_from = metric,
              values_from = part_df)

```

```{r POS_beta_div_across_sites_within_years, echo = F}

inset <- site.year.part %>% 
  pivot_longer(cols = c(`Repl/BDtotal`, `RichDif/BDtotal`),
               names_to = "type.contr",
               values_to = "contribution") %>% 
  ggplot(aes(x = year, y = contribution, fill = type.contr)) +
  geom_col() +
  theme_cowplot() +
  scale_fill_discrete(labels = c("species\nturnover",
                                 "richness\ndifferences"),
                      name = "") +
  labs(x = "", y = "Contribution to\nBeta Diversity") +
  theme(axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8, angle = 45),
        axis.title = element_text(size = 10))

plot.y <- site.year.part %>% 
  mutate(year = as.numeric(year) + 2013) %>% 
  ggplot(aes(x = year, y = BDtotal)) +
  geom_line() +
  theme_cowplot() +
  ylim(.24, .40) +
  labs(x = "Year", y = "Beta diversity")

ggdraw(plot.y) +
  draw_plot(inset, .25, .49, .7, .45)

```


# What is the turnover of species within major sites across years?

Can we answer this question if the sampling was not the same across major sites?






# How similar are populations within sites based on distance (spatial autocorrelation), and can we assess this for particular bee groups

## Interesting groups: Osmia (lots of species, not abundant), Halictus

# Can we assess why groups like Halictus have so much variability in abundance, and whether this has anything to do with the sex of the individuals collected?

# Is there greater overall diversity at SCL vs POS vs PBF?

# Distance effects – can we estimate flight

# Can we compare to Eli’s study?

# What species are persistent?


## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


