---
title: "BeeSearch initial"
author: "Dr. Riley M. Anderson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = F,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

What is this analysis about?

### Summary of Results
* 

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(vegan)
library(adespatial)
library(knitr)
library(fossil)
library(geosphere)
library(randomForest)
library(caret)


# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
dat1 <- read.csv("data/CABS1.csv")

comp.sites <- read.csv("data/comparable.sites.csv")

site.locs <- read.csv("data/site.locations.csv")
####################################
# Funtions
####################################

source("functions/TES.R")


```


```{r Data_Wrangling, echo = F, comment = ""}

# general cleaning and conversion:
dat1 <- dat1 %>% 
  mutate(across(c(Station : Sex, Site), as.factor),
         Date = ymd(Label.Date),
         Year = year(Date),
         Month = month(Date),
         # JD --- Day of year (julian day)
         JD = yday(Date),
         # ToY --- Time of Year
         ToY = case_when(JD < 135 ~ "early",
                         JD > 205 ~ "late",
                         TRUE ~ "mid"),
         StationYear = paste(Station, Year, sep = "_")) %>% 
  filter(Station != "") %>% 
  mutate(Sex = case_when(
      Sex == "M" ~ "male",
      Sex == "F" ~ "female",
      TRUE ~ NA_character_ # Handle any unexpected values
    ))

incomp.sites <- dat1 %>% 
  select(StationYear) %>% 
  filter(StationYear == "POS1_2018" |
           StationYear == "POS10_2019" |
           StationYear == "POS11_2018" |
           StationYear == "POS13_2017" |
           StationYear == "POS14_2017" |
           StationYear == "POS15_2017" |
           StationYear == "POS16_2017" |
           StationYear == "POS17_2017" |
           StationYear == "POS18_2017" |
           StationYear == "POS19_2017" |
           StationYear == "POS20_2017" |
           StationYear == "POS20_2019" |
           StationYear == "POS21_2017" |
           StationYear == "POS22_2019" |
           StationYear == "POS23_2019") %>% 
  distinct() %>% 
  mutate(StationYear = factor(StationYear))


# Sites with comparable sampling:
comp.sites <- comp.sites %>% 
  mutate(StationYear = paste(Station, Year, sep = "_"))

# filter out the sites with incompatible sampling efforts:
comp.data <- semi_join(dat1, comp.sites, by = "StationYear")

dat2 <- anti_join(dat1, incomp.sites, by = "StationYear") %>% 
  filter(Year != 2014 & Year != 2020) %>% 
  mutate(Year = factor(Year))

comp.data <- comp.data %>% 
  mutate(across(c(Year, StationYear), as.factor))


######################################################
# pairwise distances

# Create a data frame with all combinations of stations
combinations <- expand.grid(Station1 = site.locs$Station,
                            Station2 = site.locs$Station)

# Merge to get lat/long for both stations in each combination
distances <- combinations %>%
  left_join(site.locs, by = c("Station1" = "Station")) %>%
  left_join(site.locs, by = c("Station2" = "Station"),
            suffix = c("_1", "_2"))

# Calculate the pairwise distances
distances <- distances %>%
  mutate(distance = distHaversine(cbind(long_1, lat_1),
                                  cbind(long_2, lat_2)))

distances <- distances %>%
  rowwise() %>%
  mutate(
    Station1 = min(as.character(Station1), as.character(Station2)),
    Station2 = max(as.character(Station1), as.character(Station2))
  ) %>%
  ungroup()

# Remove duplicate rows and selct relevant columns
distances <- distances %>%
  distinct(Station1, Station2, .keep_all = TRUE) %>% 
  select(Station1, Station2, distance) %>% 
  filter(distance != 0) %>% 
  mutate(temp.id = case_when(
    Station1 == Station2 ~ 0,
    TRUE ~ 1
  )) %>% 
  filter(temp.id == 1) %>% 
  select(-temp.id)

dist.matrix <- distances %>% 
  pivot_wider(names_from = Station2,
              values_from = distance,
              values_fill = list(distance = 0))



##################################################
# matrix data

## matrices by year:

get_matrix_sites_within_years <- function(input_year, data){
  matrix.site.years <- data %>% 
  filter(Year == input_year) %>% 
  group_by(Station, Short.label.name) %>% 
  summarise(n = sum(Count)) %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n,
              values_fill = list(n = 0))  %>% 
  column_to_rownames("Station")
  
  return(matrix.site.years)
}


## matrices by year and time of season:
mat2017 <- dat2 %>% 
  filter(Year == "2017" & Site == "POS" &
           Collection.Method == "T") %>% 
  group_by(Station, Short.label.name, ToY) %>%
  summarise(n = sum(Count)) %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n) %>% 
  replace(is.na(.), 0) %>% 
  mutate(StationTime = paste(Station, ToY, sep = "_")) %>% 
  column_to_rownames("StationTime") %>% 
  select(-Station, -ToY)

meta2017 <- dat2 %>% 
  filter(Year == "2017" & Site == "POS" &
           Collection.Method == "T") %>% 
  select(Station, ToY) %>% 
  distinct() %>% 
  mutate(StationTime = paste(Station, ToY, sep = "_")) %>% 
  column_to_rownames("StationTime")


## matrix by time of season:
mat.seasons <- dat2 %>% 
  filter(Collection.Method == "T" &
           Site == "POS") %>% 
  group_by(Station, Short.label.name, ToY, Year) %>%
  summarise(n = sum(Count)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = n) %>% 
  replace(is.na(.), 0) %>% 
  mutate(StationTimeYear = paste(Station, ToY, Year, sep = "_")) %>% 
  column_to_rownames("StationTimeYear") %>% 
  select(-Station, -ToY, -Year)

meta.seasons <- dat2 %>% 
  filter(Site == "POS" &
           Collection.Method == "T") %>% 
  select(Station, ToY, Year) %>% 
  distinct() %>% 
  mutate(StationTimeYear = paste(Station, ToY, Year, sep = "_")) %>% 
  column_to_rownames("StationTimeYear")


```

# How many species do we have at each location? How much variability is there in richness?

```{r Q1, echo = F}

rich.year <- comp.data %>% 
  select(Year, Short.label.name) %>% 
  group_by(Year) %>% 
  distinct() %>% 
  tally()

rich.site <- comp.data %>% 
  select(Station, Short.label.name) %>% 
  group_by(Station) %>% 
  distinct() %>% 
  tally()

rich.year.site <- comp.data %>% 
  select(StationYear, Short.label.name) %>% 
  group_by(StationYear) %>% 
  distinct() %>% 
  tally()

rich.site.year.sep <- comp.data %>% 
  select(Station, Year, Short.label.name) %>% 
  group_by(Station, Year) %>% 
  distinct() %>% 
  tally()

rich.site.year.sep %>% 
  filter(Year != "2014" & Year != "2020") %>% 
  ggplot(aes(x = Year, y = n, color = Station, group = Station)) +
  geom_point() +
  geom_line() +
  labs(y = "Richness") +
  theme_classic()

```
This is purely species counts, no kind of offset for sampling effort (other than excluding 2014 and 2020).



# Sex ratio in trap vs. netting collection method:
```{r sex_ratios_function, echo = F}

sex_ratio_function <- function(data, group){
  data  %>%
    group_by(across(all_of(group))) %>%
    tally(name = "count") %>%
    pivot_wider(names_from = Sex,
                values_from = count,
                values_fill = list(count = 0)) %>%
    mutate(sex_ratio = male/female)
}

```
* Ratios > 1 represent male bias
* Ratios < 1 represent female bias


* Overall:
```{r overall_sex_ratios, echo = F}
kable(sex_ratio_function(dat2, c("Sex", "Collection.Method")),
      format = "markdown", digits = 2)
```

* By site:
```{r sex_ratios_by_site, echo = F}
kable(sex_ratio_function(dat2, c("Sex", "Collection.Method", "Site")),
      format = "markdown", digits = 2)
```

* By season:
```{r sex_ratios_by_season, echo = F}
kable(sex_ratio_function(dat2, c("Sex", "Collection.Method", "ToY")),
      format = "markdown", digits = 2)
```

* By season and site:
```{r sex_ratios_by_season_and_site, echo = F}
kable(sex_ratio_function(dat2, c("Sex", "Collection.Method",
                                 "ToY", "Site")),
      format = "markdown", digits = 2)
```

# Chao indices
```{r chao_indices, echo = F, cache = F}

# Sites by season
chao_site_season <- dat2 %>% 
  filter(Collection.Method == "T") %>% 
  unite("SiteSeason", c(Site, ToY), sep = "_") %>% 
  group_by(SiteSeason, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = SiteSeason,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Short.label.name")

chao_site_season <- data.frame(
  chao1 = sapply(seq_along(chao_site_season), function(i){
    chao1(chao_site_season[[i]])}),
  SiteSeason = colnames(chao_site_season)) %>% 
  separate_wider_delim(SiteSeason, delim = "_",
                       names = c("Site", "Season"))

# Sites
chao_site <- dat2 %>% 
  filter(Collection.Method == "T") %>% 
  group_by(Site, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = Site,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Short.label.name")

chao_site <- data.frame(
  chao1 = sapply(seq_along(chao_site), function(i){
    chao1(chao_site[[i]])}),
  Site = colnames(chao_site))

# Stations

chao_stations <- dat2 %>% 
  filter(Collection.Method == "T") %>% 
  group_by(Station, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = Station,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Short.label.name")

chao_stations <- data.frame(
  chao1 = sapply(seq_along(chao_stations), function(i){
    chao1(chao_stations[[i]])}),
  Station = colnames(chao_stations))


# Station by year

chao_station_year <- dat2 %>% 
  filter(Collection.Method == "T") %>% 
  group_by(StationYear, Short.label.name) %>% 
  tally(name = "count") %>% 
  pivot_wider(names_from = StationYear,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Short.label.name")

chao_station_year <- data.frame(
  chao1 = sapply(seq_along(chao_station_year), function(i){
    chao1(chao_station_year[[i]])}),
  StationYear = colnames(chao_station_year)) %>% 
  separate_wider_delim(StationYear, delim = "_",
                       names = c("Station", "Year"))
```
From Anne Chao 1989:

Chao1 minimum species richness is defined non-parametrically as:

$$S_{est} = S_{obs} + f_{1}^{2}/(2f_{2})$$


* Chao indices by site
```{r chao_site, echo = F}
kable(chao_site, format = "markdown", digits = 2)
```

* Chao indices by site and season
```{r chao_site_season, echo = F, cache = F}
kable(chao_site_season, format = "markdown", digits = 2)
```

* Chao indices by station
```{r chao_station, echo = F}
kable(chao_stations, format = "markdown", digits = 2)
```

* Chao indices by station and year
```{r chao_station_year, echo = F}
kable(chao_station_year, format = "markdown", digits = 2)

chao_station_year %>% 
  mutate(Year = as.numeric(Year),
         Site = substr(Station, 1,3)) %>% 
  ggplot(aes(x = Year, y = chao1, color = Site)) +
  geom_jitter(shape = 21, width = 0.05) +
  geom_smooth(aes(fill = Site),
              method = "lm", alpha = 0.25) +
  theme_cowplot() +
  labs(y = "Chao1\n(minimum species richness)")

```
** Minimum species richness (Chao1) across time.** Points are stations (sub sites), lines are fit by linear regression, and the shaded regions are 95% confidence intervals. The data exclude 2014 and 2020, and all net caught records.



# Does community composition change over time?


# How does community composition change across seasons?
```{r composition_across_seasons, echo = F, cache=F}

adonis2(mat.seasons ~ ToY * Year,
        data = meta.seasons, method = "bray")

permutest(betadisper(vegdist(mat.seasons, method = "bray"),
    meta.seasons$Year))


nmds.seasons <- metaMDS(mat.seasons, distance = "bray", k = 2,
                        autotransform = F, try = 250)

# stressplot(nmds.seasons)
nmds.seasons$stress

# site scores
site_scores_seasons <- as.data.frame(scores(nmds.seasons)$sites)

site_scores_seasons <- merge(site_scores_seasons,
                        select(meta.seasons, c(Station, ToY, Year)),
                        by.x = "row.names", by.y = "row.names")

# species scores
spp_scores_seasons <- as.data.frame(scores(nmds.seasons)$species)

spp_scores_seasons <- cbind(spp_scores_seasons,
                       Species = row.names(spp_scores_seasons))
```

# What bee groups drive changes in abundance from early to mid to late year?
```{r nmds_fig_seasons, echo = F, cache=F}

site_scores_seasons <- site_scores_seasons %>%  
  mutate(ToY = factor(ToY),
         Year = factor(Year)) %>% 
  as.data.frame()


spp_scores_seasons <- as.data.frame(spp_scores_seasons)

centroid_seasons <- site_scores_seasons %>% 
  group_by(ToY) %>% 
  summarise(NMDS1  = mean(NMDS1), NMDS2 = mean(NMDS2))

site_scores_seasons %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = ToY)) +
  geom_point(aes(NMDS1, NMDS2, color = ToY),
             size = 2) +
  stat_ellipse() +
  geom_point(data = centroid_seasons, size = 5, shape = 21,
             color = 'black', aes(fill = ToY), show.legend = F) +
  labs(color = "Time of Season") +
  ggrepel::geom_text_repel(data = spp_scores_seasons,
                           aes(x = NMDS1, y = NMDS2, label = Species),
                           size = 1.5, inherit.aes = FALSE,
                           max.overlaps = 20) +
  theme_cowplot()


```


# How does community composition change over time?
```{r nmds_fig_years, echo = F, cache=F}

centroid_years <- site_scores_seasons %>% 
  group_by(Year) %>% 
  summarise(NMDS1  = mean(NMDS1), NMDS2 = mean(NMDS2))

site_scores_seasons %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = Year)) +
  geom_point(aes(NMDS1, NMDS2, color = Year),
             size = 2) +
  stat_ellipse() +
  geom_point(data = centroid_years, size = 5, shape = 21,
             color = 'black', aes(fill = Year), show.legend = F) +
  labs(color = "Year") +
  ggrepel::geom_text_repel(data = spp_scores_seasons,
                           aes(x = NMDS1, y = NMDS2, label = Species),
                           size = 1.5, inherit.aes = FALSE,
                           max.overlaps = 20) +
  theme_cowplot()

```


# All sites NMDS

* By Site/Year
```{r nmds_all_sites, echo = F}

#####################################################
# matrix assembly
site_matrix <- dat2 %>% 
  filter(Collection.Method == "T") %>% 
  group_by(Short.label.name, StationYear) %>% 
  tally(n = "count") %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("StationYear")

site_meta <- dat2 %>% 
  filter(Collection.Method == "T") %>% 
  select(Site, Station, StationYear, Year) %>% 
  distinct()
#####################################################



#############################################################
# Stats
adonis2(site_matrix ~ Site + Year,
        data = site_meta, method = "bray")

permutest(betadisper(vegdist(site_matrix, method = "bray"),
                     site_meta$Year))

permutest(betadisper(vegdist(site_matrix, method = "bray"),
                     site_meta$Site))
#############################################################


########################################################
# NMDS scores
nmds_SSY <- metaMDS(site_matrix, distance = "bray",
                    k = 3, try = 250, autotransform = F)
nmds_SSY$stress
# 
# stressplot(nmds_SSY)

site_scores <- data.frame(scores(nmds_SSY)$sites) %>%
  rownames_to_column("StationYear") %>% 
  left_join(site_meta, by = "StationYear") %>%
  unite(SiteYear, c("Site", "Year"), sep = "_", remove = F) %>% 
  mutate(across(c(SiteYear:Year), as.factor)) %>% 
  filter(Site != "") %>% 
  droplevels()

spp_scores <- data.frame(scores(nmds_SSY)$species) %>% 
  rownames_to_column("Species")

centroids_SSY <- site_scores %>% 
  group_by(SiteYear) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))

site_scores %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = SiteYear)) +
  geom_point(aes(NMDS1, NMDS2, color = SiteYear),
             size = 2) +
  stat_ellipse() +
  theme_cowplot()

```

## NMDS by Site
```{r nmds_all_sites_year, echo = F}
site_scores %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = Site)) +
  geom_point(aes(NMDS1, NMDS2, color = Site),
             size = 2) +
  stat_ellipse() +
  theme_cowplot() +
  ggrepel::geom_text_repel(aes(label = Station))
```


### Site classification by species composition (Random Forest)
```{r random_forest_sites, echo = F}

rf_matrix <- site_matrix %>% 
  rownames_to_column("StationYear") %>% 
  separate_wider_delim(StationYear, delim = "_",
                       cols_remove = F,
                       names = c("Station", "Year")) %>% 
  separate_wider_position(Station,
                          widths = c(Site = 3, 3, StationID = 2),
                          cols_remove = F,
                          too_few = "debug") %>% 
  select(-StationID, -Station_width,
         -Station_remainder, -Station_ok) %>% 
  mutate(across(c(Site:StationYear), as.factor))

rfmod_site <- randomForest(y = rf_matrix$Site,
                           x = select(rf_matrix,
                                      -Year, -Station,
                                      -StationYear, -Site),
                           importance = T, nPerm = 999)


train(y = rf_matrix$Site,
      x = select(rf_matrix,
                 -Year, -Station,
                 -StationYear, -Site),
      method = "rf", na.action = na.omit)

update(rfmod_site, mtry = 73)

plot(rfmod_site)

imp_vals_top <- data.frame(importance(rfmod_site)) %>% 
  arrange(desc(MeanDecreaseGini)) %>% 
  filter(MeanDecreaseGini > quantile(MeanDecreaseGini, 0.9))

kable(imp_vals_top, format = "markdown", digits = 3)

```
** Random Forest classification of site by species composition.** The model was tuned without pre-processing. Bootstrapped resampling used 25 replicates. Overall model accuracy was 76.4%. The model clearly delineated the BPF sites with 0.00 class error. Similarly, POS sites were near perfect with 0.029 class error. However, the SCL sites were never classified correctly (1.00 class error), with most (8/9) identified as POS. This lends further support to the NMDS figure above, species composition is very similar across SCL and POS, but BPF sites have different community composition compared to the other two sites.

The table above shows the species most representative of the community differences used to make the above classifications. These species are the top 10th percentile of ranked variable importance (mean decrease in Gini score). Take this with a grain of salt, the model didn't do a great job, overall Out of Bag error rate = 19.3%. This is almost unacceptable, especially because Random Forest more often overfits the data.



* By year
```{r nmds_all_year, echo = F}
site_scores %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = Year)) +
  geom_point(aes(NMDS1, NMDS2, color = Year),
             size = 2) +
  stat_ellipse() +
  theme_cowplot()
```

# POS 2017 Changes over seasons

```{r NMDS_2017_stress, include = F}

nmds2017 <- metaMDS(mat2017, distance = "bray",
                      k = 2, try = 250, autotransform = F)


nmds2017$dist
stressplot(nmds2017)

```

```{r nmds2017_meta_merge, echo = F}

nmds2017dist <- mat2017 %>% 
  avgdist(sample = 5)


# site scores
site_scores2017 <- as.data.frame(scores(nmds2017)$sites)

site_scores2017 <- merge(site_scores2017,
                        select(meta2017, c(Station, ToY)),
                        by.x = "row.names", by.y = "row.names")

# species scores
spp_scores2017 <- as.data.frame(scores(nmds2017)$species)

spp_scores2017 <- cbind(spp_scores2017,
                       Species = row.names(spp_scores2017))


adonis2(nmds2017dist ~ meta2017$ToY, method = "bray")
permutest(betadisper(nmds2017dist, meta2017$ToY))


```

Within the Port of Seattle sites in 2017 and only 2017, species composition changes significantly throughout the season with distinct groups of species in early, mid, and late season sampling.

```{r nmds2017_plot, echo = F}
site_scores2017 <- site_scores2017 %>%  
  mutate(ToY = factor(ToY)) %>% 
  as.data.frame()


spp_scores2017 <- as.data.frame(spp_scores2017)

centroid2017 <- site_scores2017 %>% 
  group_by(ToY) %>% 
  summarise(NMDS1  = mean(NMDS1), NMDS2 = mean(NMDS2))

site_scores2017 %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = ToY)) +
  geom_point(aes(NMDS1, NMDS2, color = ToY),
             size = 2) +
  stat_ellipse() +
  geom_point(data = centroid2017, size = 5, shape = 21,
             color = 'black', aes(fill = ToY), show.legend = F) +
  labs(color = "Time of Season") +
  ggrepel::geom_text_repel(data = spp_scores2017,
                           aes(x = NMDS1, y = NMDS2, label = Species),
                           size = 1.5, inherit.aes = FALSE,
                           max.overlaps = 20) +
  theme_cowplot()


```
**Non-metric multidimensional scaling of bee species in Port of Seattle in 2017.** Points are sub-sites within the port of Seattle in 2017. They are separated by time of season with the 8 sub sites at the early season in red, the same 8 sub sites at the middle season in blue, and the same 8 sub sites at the end of the season in green. The NMDS space represents total bee species composition and the labels are specific bee species and their position in NMDS space. Large points are the centroids (means) of the points in NMDS space. Ellipses are 95% confidence intervals around the centroids. 

# Can we do analysis with and without major species (H. tripartitus; A. texanus; M. microsticta/us)

What analysis?


# What is the turnover of species within sub-sites within years?

```{r species_turnover_site_years_POS, echo = F}


matrix.site.years.list <- lapply(levels(comp.data$Year),
                                 data = comp.data,
                                 get_matrix_sites_within_years)

bdiv.list <- lapply(seq_along(matrix.site.years.list), function(i){
  beta.div.comp(matrix.site.years.list[[i]],
                coef = "J", quant = T)
})

extract_and_combine <- function(list_of_lists, part_name) {
  # Extract the specified part from each sublist and combine into a single data frame
  combined_df <- map2_df(list_of_lists, seq_along(list_of_lists), ~ {
    part_df <- .x[[part_name]]
    part_df <- as.data.frame(part_df)
    part_df <- part_df %>%
      mutate(ListID = .y)  # Add a unique identifier column
    
    return(part_df)
  })
  
  return(combined_df)
}

site.year.part <- extract_and_combine(bdiv.list, "part")

site.year.part <- site.year.part %>% 
  mutate(year = factor(ListID + 2013)) %>% 
  rownames_to_column(var = "metric") %>% 
  mutate(metric = str_remove(metric, "\\.\\.\\..*$")) %>% 
  select(-ListID) %>% 
  pivot_wider(names_from = metric,
              values_from = part_df)

```

### POS
```{r POS_beta_div_across_sites_within_years, echo = F}

inset <- site.year.part %>% 
  pivot_longer(cols = c(`Repl/BDtotal`, `RichDif/BDtotal`),
               names_to = "type.contr",
               values_to = "contribution") %>% 
  ggplot(aes(x = year, y = contribution, fill = type.contr)) +
  geom_col() +
  theme_cowplot() +
  scale_fill_discrete(labels = c("species\nturnover",
                                 "richness\ndifferences"),
                      name = "") +
  labs(x = "", y = "Contribution to\nBeta Diversity") +
  theme(axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8, angle = 45),
        axis.title = element_text(size = 10))

plot.y <- site.year.part %>% 
  mutate(year = as.numeric(year) + 2013) %>% 
  ggplot(aes(x = year, y = BDtotal)) +
  geom_line() +
  theme_cowplot() +
  ylim(.24, .40) +
  labs(x = "Year", y = "Beta diversity")

ggdraw(plot.y) +
  draw_plot(inset, .25, .49, .7, .45)

```

### SCL
```{r species_turnover_site_years_SCL, echo = F}

SCL <- dat2 %>% 
  filter(Site == "SCL") %>% 
  droplevels()

scl.matrix.list <- lapply(levels(SCL$Year),
                                 data = SCL,
                                 get_matrix_sites_within_years)


bdiv.list.scl <- lapply(seq_along(scl.matrix.list), function(i){
  beta.div.comp(scl.matrix.list[[i]],
                coef = "J", quant = T)
})

site.year.part.scl <- extract_and_combine(bdiv.list.scl, "part")

site.year.part.scl <- site.year.part.scl %>% 
  mutate(year = factor(ListID + 2014)) %>% 
  rownames_to_column(var = "metric") %>% 
  mutate(metric = str_remove(metric, "\\.\\.\\..*$")) %>% 
  select(-ListID) %>% 
  pivot_wider(names_from = metric,
              values_from = part_df)

```

```{r SCL_beta_div_across_sites_within_years, echo = F}

inset.scl <- site.year.part.scl %>% 
  pivot_longer(cols = c(`Repl/BDtotal`, `RichDif/BDtotal`),
               names_to = "type.contr",
               values_to = "contribution") %>% 
  ggplot(aes(x = year, y = contribution, fill = type.contr)) +
  geom_col() +
  theme_cowplot() +
  scale_fill_discrete(labels = c("species\nturnover",
                                 "richness\ndifferences"),
                      name = "") +
  labs(x = "", y = "Contribution to\nBeta Diversity") +
  theme(axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8, angle = 45),
        axis.title = element_text(size = 10))

plot.y.scl <- site.year.part.scl %>% 
  mutate(year = as.numeric(year) + 2014) %>% 
  ggplot(aes(x = year, y = BDtotal)) +
  geom_line() +
  theme_cowplot() +
  ylim(.27, .45) +
  labs(x = "Year", y = "Beta diversity") +
  scale_x_continuous(breaks = c(2015, 2016))

ggdraw(plot.y.scl) +
  draw_plot(inset.scl, .25, .49, .7, .45)

```


### BPF
```{r species_turnover_site_years_BPF, echo = F}
BPF <- dat2 %>% 
  filter(Site == "BPF") %>% 
  droplevels()

BPF.matrix.list <- lapply(levels(BPF$Year),
                          data = BPF,
                          get_matrix_sites_within_years)

bdiv.list.bpf <- lapply(seq_along(BPF.matrix.list), function(i){
  beta.div.comp(BPF.matrix.list[[i]],
                coef = "J", quant = T)
})

site.year.part.bpf <- extract_and_combine(bdiv.list.bpf, "part")

site.year.part.bpf <- site.year.part.bpf %>% 
  mutate(year = factor(ListID + 2017)) %>% 
  rownames_to_column(var = "metric") %>% 
  mutate(metric = str_remove(metric, "\\.\\.\\..*$")) %>% 
  select(-ListID) %>% 
  pivot_wider(names_from = metric,
              values_from = part_df)

```

```{r BPF_beta_div_across_sites_within_years, echo = F}

inset.bpf <- site.year.part.bpf %>% 
  pivot_longer(cols = c(`Repl/BDtotal`, `RichDif/BDtotal`),
               names_to = "type.contr",
               values_to = "contribution") %>% 
  ggplot(aes(x = year, y = contribution, fill = type.contr)) +
  geom_col() +
  theme_cowplot() +
  scale_fill_discrete(labels = c("species\nturnover",
                                 "richness\ndifferences"),
                      name = "") +
  labs(x = "", y = "Contribution to\nBeta Diversity") +
  theme(axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8, angle = 45),
        axis.title = element_text(size = 10))

plot.y.bpf <- site.year.part.bpf %>% 
  mutate(year = as.numeric(year) + 2017) %>% 
  ggplot(aes(x = year, y = BDtotal)) +
  geom_line() +
  theme_cowplot() +
  ylim(.33, .50) +
  labs(x = "Year", y = "Beta diversity") +
  scale_x_continuous(breaks = c(2018, 2019))

ggdraw(plot.y.bpf) +
  draw_plot(inset.bpf, .25, .49, .7, .45)

```



# Is there greater overall diversity at SCL vs POS vs PBF?
```{r diversity_among_sites, echo = F}
# Site level matrix
site_matrix <- dat1 %>% 
  group_by(Short.label.name, Site) %>% 
  tally(n = "count") %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("Site")

h <- diversity(site_matrix) %>% 
  enframe(name = "Site", value = "Shannon")

simp <- diversity(site_matrix, index = "simpson") %>% 
  enframe(name = "Site", value = "Simpson") %>% 
  select(-Site)

invsimp <- diversity(site_matrix, index = "inv") %>% 
  enframe(name = "Site", value = "InvSimpson") %>% 
  select(-Site)

unbias.simp <- simpson.unb(site_matrix) %>% 
  enframe(name = "Site", value = "UnbiasedSimpson") %>% 
  select(-Site)

f.alpha <- fisher.alpha(site_matrix) %>% 
  enframe(name = "Site", value = "FisherAlpha") %>% 
  select(-Site)



kable(bind_cols(h, simp, invsimp, unbias.simp, f.alpha),
      format = "markdown", digits = 2)
```
Diversity is similar across all 3 sites.



## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


