---
title: "BeeSearch Multivariate Analyses"
author: "Dr. Riley M. Anderson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = F,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

What is this analysis about?

### Summary of Results
* 

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(vegan)
library(randomForest)
library(caret)
library(knitr)



# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
dat1 <- read.csv("data/CABS1.csv")

####################################

###################################
###################
# color palette
c24 <- c(
  "maroon", "#E31A1C", "green4", "#6A3D9A",
  "#FF7F00", "gold1", "mediumpurple1","skyblue2", "#FB9A99",
  "palegreen2", "#CAB2D6", "#FDBF6F", "gray70", "khaki2",
  "dodgerblue2", "orchid1", "deeppink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3", "darkorange4"
)


```


```{r Data_Wrangling, echo = F, comment = ""}

# general cleaning and conversion:
dat1 <- dat1 %>% 
  uncount(Count) %>% 
  mutate(across(c(Station : Sex, Site), as.factor), # convert to factors
         Date = ymd(Label.Date), # Convert to date
         Year = year(Date), # Create year column
         Month = month(Date), # create month column
         # JD --- Day of year (julian day)
         JD = yday(Date),
         # ToY --- Time of Year
         ToY = case_when(JD < 135 ~ "early",
                         JD > 205 ~ "late",
                         TRUE ~ "mid"),
         # Create a station/year column for grouping
         StationYear = paste(Station, Year, sep = "_")) %>% 
  filter(Station != "",
         Site != "") %>% # remove empty station & site records
  # convert M/F to male/female to avoid "F = FALSE" misinterpretation
  mutate(Sex = case_when(
      Sex == "M" ~ "male",
      Sex == "F" ~ "female",
      TRUE ~ NA_character_ # Handle any unexpected values
    )) %>% 
  # correct misidentification record
  mutate(Short.label.name = case_when(
    Short.label.name == "Megachile apicalis" ~ "Megachile rotundata",
    TRUE ~ as.character(Short.label.name)
  )) %>% 
  # correct misindentification sex classification
  mutate(Sex = case_when(
    Unique.Specimen.Number == 678 |
      Unique.Specimen.Number == 679 ~ "Male",
    TRUE ~ as.character(Sex)
  ))


# keep the single record of morphospecies "Hylaeus sp. 2023_1"
keep.Hylaeus <- dat1 %>% 
  filter(str_detect(Short.label.name, "Hylaeus sp. 2023"))

# remove any records where species is undetermined:
dat2 <- dat1 %>% 
  filter(!str_detect(Short.label.name, "(?<!m)sp\\."),
         !str_detect(Short.label.name, "(?<!m)msp\\.")) %>% 
  bind_rows(keep.Hylaeus)


```


## Across sites
```{r nmds_all_sites_prep_and_tests, echo = F}

#####################################################
# matrix assembly
site_matrix <- dat2 %>% 
  filter(Collection.Method == "T") %>% 
  group_by(Short.label.name, StationYear) %>% 
  tally(n = "count") %>% 
  pivot_wider(names_from = Short.label.name,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("StationYear")

site_meta <- dat2 %>% 
  filter(Collection.Method == "T") %>% 
  select(Site, Station, StationYear, Year) %>% 
  distinct()
#####################################################



#############################################################
# Stats
site.year.mod <- adonis2(site_matrix ~ Site * Year,
        data = site_meta, method = "bray")

site.year.mod

permutest(betadisper(vegdist(site_matrix, method = "bray"),
                     site_meta$Year))

permutest(betadisper(vegdist(site_matrix, method = "bray"),
                     site_meta$Site))

```


```{r nmds_sites_calcs, eval = T, include = F}
# NMDS scores
nmds_SSY <- metaMDS(site_matrix, distance = "bray",
                    k = 3, try = 250, autotransform = F)
nmds_SSY$stress
# 
# stressplot(nmds_SSY)

site_scores <- data.frame(scores(nmds_SSY)$sites) %>%
  rownames_to_column("StationYear") %>% 
  left_join(site_meta, by = "StationYear") %>%
  unite(SiteYear, c("Site", "Year"), sep = "_", remove = F) %>% 
  mutate(across(c(SiteYear:Year), as.factor)) %>% 
  filter(Site != "") %>% 
  droplevels()

spp_scores <- data.frame(scores(nmds_SSY)$species) %>% 
  rownames_to_column("Species")

centroids_SSY <- site_scores %>% 
  group_by(SiteYear) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))

centroids_site <- site_scores %>% 
  group_by(Site) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))

centroids_year <- site_scores %>% 
  group_by(Year) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))
```

### Site classification by species composition (Random Forest)
```{r random_forest_sites, echo = F}

rf_matrix <- site_matrix %>% 
  rownames_to_column("StationYear") %>% 
  separate_wider_delim(StationYear, delim = "_",
                       cols_remove = F,
                       names = c("Station", "Year")) %>% 
  separate_wider_position(Station,
                          widths = c(Site = 3, 3, StationID = 2),
                          cols_remove = F,
                          too_few = "debug") %>% 
  select(-StationID, -Station_width,
         -Station_remainder, -Station_ok) %>% 
  mutate(across(c(Site:StationYear), as.factor))

rfmod_site <- randomForest(y = rf_matrix$Site,
                           x = select(rf_matrix,
                                      -Year, -Station,
                                      -StationYear, -Site),
                           importance = T, nPerm = 999,
                           proximity = T)


train(y = rf_matrix$Site,
      x = select(rf_matrix,
                 -Year, -Station,
                 -StationYear, -Site),
      method = "rf", na.action = na.omit)

update(rfmod_site, mtry = 55)


imp_vals_top <- data.frame(importance(rfmod_site)) %>% 
  arrange(desc(MeanDecreaseGini)) %>% 
  filter(MeanDecreaseGini > quantile(MeanDecreaseGini, 0.9))

kable(imp_vals_top, format = "markdown", digits = 3)


dat2 %>% filter(Short.label.name == "Agapostemon texanus") %>%
  group_by(Site, StationYear) %>% tally() %>% 
  group_by(Site) %>% 
  summarise(mean = mean(n))

MDSplot(rfmod_site, rf_matrix$Site)
```

**Random Forest classification of site by species composition.** The model was tuned without pre-processing. Bootstrapped resampling used 25 replicates. The model was built with 500 trees and model tuning maximized accuracy at 55 variables/split. Overall model accuracy was 82%. Out of bag estimate of error rate was 16.7%. This lends further support to the NMDS figure below, species composition is very similar across SCL and POS, but BPF sites have different community composition compared to the other two sites. 

The table above shows the species most representative of the community differences used to make the above classifications. These species are the top 10th percentile of ranked variable importance (mean decrease in Gini score, a measure of the total decrease in node impurities from splitting on the variable, averaged over all trees).


### NMDS by Site figure (include this)
```{r nmds_all_sites_year, echo = F}

imp_spp <- imp_vals_top %>% 
  rownames_to_column("Species") %>% 
  select(Species, MeanDecreaseGini)

spp_scores <- right_join(spp_scores, imp_spp, by = "Species")

site_scores %>% 
  ggplot(aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(NMDS1, NMDS2, color = Site),
             size = 1) +
  stat_ellipse(aes(color = Site),
               show.legend = F) +
  geom_point(data = centroids_site,
             aes(x = NMDS1, y = NMDS2, fill = Site),
             size = 4,
             shape = 21, show.legend = F) +
  theme_cowplot() +
  ggrepel::geom_text_repel(data = spp_scores,
                           aes(x = NMDS1, y = NMDS2,
                               label = Species,
                               size = MeanDecreaseGini),
                           max.overlaps = 20) +
  scale_size_continuous(
    name = "      Mean\ndecrease Gini",
    limits = c(.4, 3),
    breaks = c(0.75, 2, 3),
    range = c(1, 6)) +
  guides(color = guide_legend(override.aes = list(size = 4, shape = 16))) +
  scale_color_manual(values = c("#E69F00", "#56B4E9", "#009E73"))+
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73")) +
  labs(color = "        Site")




```
**Variation in community composition across sites.** Bee species are plotted on the first two axes of a three-dimensional non-metric multidimensional scaling ordination of the 58 combinations of station (subsite) and year, across the three sites. Small points are the individual station/year combinations. Large points are the centroids of the three sites. Ellipses are 95% confidence intervals around the site centroids.  Bees species shown are the most representative (top 10th percentile of a random forest analysis) of the compositional differences among sites. Text size of the labels is proportional to variable importance score (mean decrease in Gini score).





## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


