---
title: "BeeSearch genus plots"
author: "Dr. Riley M. Anderson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = F,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

This analysis plots the count of unique species/genus and the proportional abundance of each genus.




```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(knitr)



# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
dat1 <- read.csv("data/CABS1.csv")

comp.sites <- read.csv("data/comparable.sites.csv")

site.locs <- read.csv("data/site.locations.csv")


####################################

```


```{r Data_Wrangling, echo = F, comment = ""}

# general cleaning and conversion:
dat1 <- dat1 %>% 
  uncount(Count) %>% 
  mutate(across(c(Station : Sex, Site), as.factor), # convert to factors
         Date = ymd(Label.Date), # Convert to date
         Year = year(Date), # Create year column
         Month = month(Date), # create month column
         # JD --- Day of year (julian day)
         JD = yday(Date),
         # ToY --- Time of Year
         ToY = case_when(JD < 135 ~ "early",
                         JD > 205 ~ "late",
                         TRUE ~ "mid"),
         # Create a station/year column for grouping
         StationYear = paste(Station, Year, sep = "_")) %>% 
  filter(Station != "",
         Site != "") %>% # remove empty station & site records
  # convert M/F to male/female to avoid "F = FALSE" misinterpretation
  mutate(Sex = case_when(
      Sex == "M" ~ "male",
      Sex == "F" ~ "female",
      TRUE ~ NA_character_ # Handle any unexpected values
    )) %>% 
  # correct misidentification record
  mutate(Short.label.name = case_when(
    Short.label.name == "Megachile apicalis" ~ "Megachile rotundata",
    TRUE ~ as.character(Short.label.name)
  )) %>% 
  # correct misindentification sex classification
  mutate(Sex = case_when(
    Unique.Specimen.Number == 678 |
      Unique.Specimen.Number == 679 ~ "Male",
    TRUE ~ as.character(Sex)
  ))


# keep the single record of morphospecies "Hylaeus sp. 2023_1"
keep.Hylaeus <- dat1 %>% 
  filter(str_detect(Short.label.name, "Hylaeus sp. 2023"))

# remove any records where species is undetermined:
dat2 <- dat1 %>% 
  filter(!str_detect(Short.label.name, "(?<!m)sp\\."),
         !str_detect(Short.label.name, "(?<!m)msp\\.")) %>% 
  bind_rows(keep.Hylaeus)







```

# Descriptive measures of genera

## All sites together
```{r genera_figs_all_sites, echo = F}

genus.counts <- dat2 %>% 
  filter(Collection.Method == "T") %>% 
  select(Genus, Short.label.name) %>% 
  group_by(Genus) %>% 
  distinct() %>% 
  tally(name = "Species")

genus.props <- dat2 %>% 
  filter(Collection.Method == "T") %>% 
  group_by(Genus) %>% 
  tally(name = "Prop") %>% 
  mutate(prop.abund = Prop/nrow(dat2))

genus <- left_join(genus.counts, genus.props, by = "Genus")

###################
# color palette
c24 <- c(
  "maroon", "#E31A1C", "green4", "#6A3D9A",
  "#FF7F00", "gold1", "mediumpurple1","skyblue2", "#FB9A99",
  "palegreen2", "#CAB2D6", "#FDBF6F", "gray70", "khaki2",
  "dodgerblue2", "orchid1", "deeppink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3", "darkorange4"
)


genus_levels <- levels(genus$Genus)

color_mapping <- setNames(c24, genus_levels)



##########
genus %>% 
  ggplot(aes(x = fct_reorder(Genus, Species, .desc = F),
             y = Species, fill = Genus)) +
  geom_col(show.legend = F) +
  geom_col(aes(y = -prop.abund * 100, fill = Genus),
           show.legend = F) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_cowplot(font_size = 16) +
  labs(x = "Genus",
       y = "") +
  annotate(geom = "text",
           x = 5,
           y = -30, 
           label = "Proportional\nabundance (%)",
           color = "black",
           size = 5) +
  annotate(geom = "text",
           x = 5,
           y = 18, 
           label = "No. unique\nspecies",
           color = "black",
           size = 5) +
  scale_y_continuous(labels = function(y) abs(y)) +
  scale_fill_manual(values = color_mapping)

#################################################
genus_plot_all <- genus %>% 
  ggplot(aes(x = fct_reorder(Genus, Species, .desc = F),
             y = Species, fill = Genus)) +
  geom_col(show.legend = F) +
  geom_col(aes(y = -prop.abund * 100, fill = Genus),
           show.legend = F) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_cowplot(font_size = 16) +
  labs(x = "",
       y = "") +
  annotate(geom = "text",
           x = 5,
           y = -30, 
           label = "Proportional\nabundance (%)",
           color = "black",
           size = 5) +
  annotate(geom = "text",
           x = 5,
           y = 18, 
           label = "Unique\nspecies",
           color = "black",
           size = 5) +
  scale_y_continuous(labels = function(y) abs(y)) +
  scale_fill_manual(values = color_mapping)

```

## POS
```{r genera_figs_pos, echo = F}

genus.counts.pos <- dat2 %>% 
  filter(Site == "POS" & 
           Collection.Method == "T") %>% 
  select(Genus, Short.label.name) %>% 
  group_by(Genus) %>% 
  distinct() %>% 
  tally(name = "Species")

genus.props.pos <- dat2 %>% 
  filter(Site == "POS" & Collection.Method == "T") %>% 
  group_by(Genus) %>% 
  tally(name = "Prop") %>% 
  mutate(prop.abund = Prop/nrow(filter(dat2, Site == "POS" &
                                         Collection.Method == "T")))

genus.pos <- left_join(genus.counts.pos, genus.props.pos, by = "Genus")

genus.pos %>% 
  ggplot(aes(x = fct_reorder(Genus, Species, .desc = F),
             y = Species, fill = Genus)) +
  geom_col(show.legend = F) +
  geom_col(aes(y = -prop.abund * 100, fill = Genus),
           show.legend = F) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_cowplot(font_size = 16) +
  labs(x = "Genus",
       y = "") +
  annotate(geom = "text",
           x = 5,
           y = -30, 
           label = "Proportional\nabundance (%)",
           color = "black",
           size = 5) +
  annotate(geom = "text",
           x = 5,
           y = 13, 
           label = "No. unique\nspecies",
           color = "black",
           size = 5) +
  scale_y_continuous(labels = function(y) abs(y)) +
  scale_fill_manual(values = color_mapping)
#################################################
genus_plot_pos <- genus.pos %>% 
  ggplot(aes(x = fct_reorder(Genus, Species, .desc = F),
             y = Species, fill = Genus)) +
  geom_col(show.legend = F) +
  geom_col(aes(y = -prop.abund * 100, fill = Genus),
           show.legend = F) +
  geom_hline(yintercept = 0) +
  scale_x_discrete(position = "top") +
  coord_flip() +
  theme_cowplot(font_size = 16) +
  labs(x = "",
       y = "") +
  annotate(geom = "text",
           x = 5,
           y = -30, 
           label = "Proportional\nabundance (%)",
           color = "black",
           size = 5) +
  annotate(geom = "text",
           x = 5,
           y = 13, 
           label = "Unique\nspecies",
           color = "black",
           size = 5) +
  scale_y_continuous(labels = function(y) abs(y)) +
  scale_fill_manual(values = color_mapping)

```

## SCL
```{r genera_figs_scl, echo = F}

genus.counts.scl <- dat2 %>% 
  filter(Site == "SCL" &
           Collection.Method == "T") %>% 
  select(Genus, Short.label.name) %>% 
  group_by(Genus) %>% 
  distinct() %>% 
  tally(name = "Species")

genus.props.scl <- dat2 %>% 
  filter(Site == "SCL" &
           Collection.Method == "T") %>% 
  group_by(Genus) %>% 
  tally(name = "Prop") %>% 
  mutate(prop.abund = Prop/nrow(filter(dat2, Site == "SCL" &
                                         Collection.Method == "T")))

genus.scl <- left_join(genus.counts.scl, genus.props.scl, by = "Genus")

genus.scl %>% 
  ggplot(aes(x = fct_reorder(Genus, Species, .desc = F),
             y = Species, fill = Genus)) +
  geom_col(show.legend = F) +
  geom_col(aes(y = -prop.abund * 100, fill = Genus),
           show.legend = F) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_cowplot(font_size = 16) +
  labs(x = "Genus",
       y = "") +
  annotate(geom = "text",
           x = 5,
           y = -15, 
           label = "Proportional\nabundance (%)",
           color = "black",
           size = 5) +
  annotate(geom = "text",
           x = 5,
           y = 10, 
           label = "No. unique\nspecies",
           color = "black",
           size = 5) +
  scale_y_continuous(labels = function(y) abs(y)) +
  scale_fill_manual(values = color_mapping)
###################################################
genus_plot_scl <- genus.scl %>% 
  ggplot(aes(x = fct_reorder(Genus, Species, .desc = F),
             y = Species, fill = Genus)) +
  geom_col(show.legend = F) +
  geom_col(aes(y = -prop.abund * 100, fill = Genus),
           show.legend = F) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_cowplot(font_size = 16) +
  labs(x = "",
       y = "") +
  annotate(geom = "text",
           x = 5,
           y = -15, 
           label = "Proportional\nabundance (%)",
           color = "black",
           size = 5) +
  annotate(geom = "text",
           x = 5,
           y = 10, 
           label = "Unique\nspecies",
           color = "black",
           size = 5) +
  scale_y_continuous(labels = function(y) abs(y)) +
  scale_fill_manual(values = color_mapping)

```

## BPF
```{r genera_figs_bpf, echo = F}

genus.counts.bpf <- dat2 %>% 
  filter(Site == "BPF" &
           Collection.Method == "T") %>% 
  select(Genus, Short.label.name) %>% 
  group_by(Genus) %>% 
  distinct() %>% 
  tally(name = "Species")

genus.props.bpf <- dat2 %>% 
  filter(Site == "BPF" & 
           Collection.Method == "T") %>% 
  group_by(Genus) %>% 
  tally(name = "Prop") %>% 
  mutate(prop.abund = Prop/nrow(filter(dat2, Site == "BPF" &
                                         Collection.Method == "T")))

genus.bpf <- left_join(genus.counts.bpf, genus.props.bpf, by = "Genus")



genus.bpf %>% 
  ggplot(aes(x = fct_reorder(Genus, Species, .desc = F),
             y = Species, fill = Genus)) +
  geom_col(show.legend = F) +
  geom_col(aes(y = -prop.abund * 100, fill = Genus),
           show.legend = F) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_cowplot(font_size = 16) +
  labs(x = "",
       y = "") +
  annotate(geom = "text",
           x = 5,
           y = -20, 
           label = "Proportional\nabundance (%)",
           color = "black",
           size = 5) +
  annotate(geom = "text",
           x = 5,
           y = 12, 
           label = "No. unique\nspecies",
           color = "black",
           size = 5) +
  scale_y_continuous(labels = function(y) abs(y)) +
  scale_fill_manual(values = color_mapping)


#########################################################
genus_plot_bpf <- genus.bpf %>% 
  ggplot(aes(x = fct_reorder(Genus, Species, .desc = F),
             y = Species, fill = Genus)) +
  geom_col(show.legend = F) +
  geom_col(aes(y = -prop.abund * 100, fill = Genus),
           show.legend = F) +
  geom_hline(yintercept = 0) +
  scale_x_discrete(position = "top") +
  coord_flip() +
  theme_cowplot(font_size = 16) +
  labs(x = "",
       y = "") +
  annotate(geom = "text",
           x = 4.5,
           y = -20, 
           label = "Proportional\nabundance (%)",
           color = "black",
           size = 5) +
  annotate(geom = "text",
           x = 4.4,
           y = 12, 
           label = "Unique\nspecies",
           color = "black",
           size = 5) +
  scale_y_continuous(labels = function(y) abs(y)) +
  scale_fill_manual(values = color_mapping)

```

## All genus plots together in one figure:
```{r genus_plot_one_fig, echo = F, fig.width=10, fig.height=10}

plot_grid(genus_plot_all, genus_plot_pos,
          genus_plot_scl, genus_plot_bpf, 
          labels = c("A", "B", "C", "D"),
          label_size = 14)

```

**Proportional abundance and unique species** A) All sites collectively, B) POS, C) SCL, and D) BPF. Data are from trap collected records and exclude all net caught records. Data are pooled across years.


## All sites together (trap and net records)
```{r genera_figs_all_sites_trap_net, echo = F}

genus.counts.tn <- dat2 %>% 
  select(Genus, Short.label.name) %>% 
  group_by(Genus) %>% 
  distinct() %>% 
  tally(name = "Species")

genus.props.tn <- dat2 %>% 
  group_by(Genus) %>% 
  tally(name = "Prop") %>% 
  mutate(prop.abund = Prop/nrow(dat2))

genus.tn <- left_join(genus.counts.tn, genus.props.tn, by = "Genus")


genus_levels.tn <- levels(genus.tn$Genus)

color_mapping.tn <- setNames(c24, genus_levels.tn)



##########
genus.tn %>% 
  ggplot(aes(x = fct_reorder(Genus, Species, .desc = F),
             y = Species, fill = Genus)) +
  geom_col(show.legend = F) +
  geom_col(aes(y = -prop.abund * 100, fill = Genus),
           show.legend = F) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_cowplot(font_size = 16) +
  labs(x = "Genus",
       y = "") +
  annotate(geom = "text",
           x = 5,
           y = -30, 
           label = "Proportional\nabundance (%)",
           color = "black",
           size = 5) +
  annotate(geom = "text",
           x = 5,
           y = 18, 
           label = "No. unique\nspecies",
           color = "black",
           size = 5) +
  scale_y_continuous(labels = function(y) abs(y)) +
  scale_fill_manual(values = color_mapping.tn)

#################################################
genus_plot_trap_net <- genus.tn %>% 
  ggplot(aes(x = fct_reorder(Genus, Species, .desc = F),
             y = Species, fill = Genus)) +
  geom_col(show.legend = F) +
  geom_col(aes(y = -prop.abund * 100, fill = Genus),
           show.legend = F) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_cowplot(font_size = 16) +
  labs(x = "",
       y = "") +
  annotate(geom = "text",
           x = 5,
           y = -30, 
           label = "Proportional\nabundance (%)",
           color = "black",
           size = 5) +
  annotate(geom = "text",
           x = 5,
           y = 18, 
           label = "Unique\nspecies",
           color = "black",
           size = 5) +
  scale_y_continuous(labels = function(y) abs(y)) +
  scale_fill_manual(values = color_mapping.tn)

```


## All sites together (net records only)
```{r genera_figs_all_sites_nets, echo = F}

genus.counts.net <- dat2 %>% 
  filter(Collection.Method == "N") %>% 
  select(Genus, Short.label.name) %>% 
  group_by(Genus) %>% 
  distinct() %>% 
  tally(name = "Species")

genus.props.net <- dat2 %>% 
  filter(Collection.Method == "N") %>% 
  group_by(Genus) %>% 
  tally(name = "Prop") %>% 
  mutate(prop.abund = Prop/nrow(dat2))

genus.net <- left_join(genus.counts.net, genus.props.net, by = "Genus")



genus_levels.net <- levels(genus.net$Genus)

color_mapping.net <- setNames(c24, genus_levels.net)



##########
genus.net %>% 
  ggplot(aes(x = fct_reorder(Genus, Species, .desc = F),
             y = Species, fill = Genus)) +
  geom_col(show.legend = F) +
  geom_col(aes(y = -prop.abund * 100, fill = Genus),
           show.legend = F) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  theme_cowplot(font_size = 16) +
  labs(x = "Genus",
       y = "") +
  annotate(geom = "text", x = 1, y = -10, label = " ") +
  annotate(geom = "text",
           x = 5,
           y = -6, 
           label = "Proportional\nabundance (%)",
           color = "black",
           size = 5) +
  annotate(geom = "text",
           x = 5,
           y = 10, 
           label = "No. unique\nspecies",
           color = "black",
           size = 5) +
  scale_y_continuous(labels = function(y) abs(y)) +
  scale_fill_manual(values = color_mapping.net)


################
genus_plot_net <- genus.net %>% 
  ggplot(aes(x = fct_reorder(Genus, Species, .desc = F),
             y = Species, fill = Genus)) +
  geom_col(show.legend = F) +
  geom_col(aes(y = -prop.abund * 100, fill = Genus),
           show.legend = F) +
  geom_hline(yintercept = 0) +
  scale_x_discrete(position = "top") +
  coord_flip() +
  theme_cowplot(font_size = 16) +
  labs(x = "",
       y = "") +
  annotate(geom = "text", x = 1, y = -10, label = " ") +
  annotate(geom = "text",
           x = 5,
           y = -6, 
           label = "Proportional\nabundance (%)",
           color = "black",
           size = 5) +
  annotate(geom = "text",
           x = 5,
           y = 10, 
           label = "No. unique\nspecies",
           color = "black",
           size = 5) +
  scale_y_continuous(labels = function(y) abs(y)) +
  scale_fill_manual(values = color_mapping.net)

```

## Genus Trap and Net, And Net Only
```{r trapnet_and_net, echo = F, fig.width = 10}

plot_grid(genus_plot_trap_net, genus_plot_net,
          labels = c("A", "B"),
          label_size = 14)

```

**Proportional abundance and unique species** for A) all net and trap caught records and B) net caught records only. Data combine records across all sites and years.


## Genus Trap only and Net only
```{r trap_and_net, echo = F, fig.width = 10}

plot_grid(genus_plot_all, genus_plot_net,
          labels = c("A", "B"),
          label_size = 14)

```

**Proportional abundance and unique species** for A) trap caught records and B) net caught records. Data combine records across all sites and years.


## All Genus plots
```{r all_plots, echo = F, fig.width = 11, fig.height = 15}

plot_grid(genus_plot_all, genus_plot_pos,
          genus_plot_scl, genus_plot_bpf,
          genus_plot_trap_net, genus_plot_net,
          labels = c("A", "B", "C", "D", "E", "F"),
          label_size = 14, nrow = 3)

```

**Proportional abundance and unique species for all genera.** A) trap caught records pooled across all sites and years, B) trap records from POS, C) trap records from SCL, D) trap records from BPF, E) trap and net caught records pooled across all sites and years, and F) all net caught records pooled across sites and years.

## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


