---
title: "BeeSearch ridge plots"
author: "Dr. Riley M. Anderson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = F,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

This analysis compares the abundance of genera and species throughout the season. All analyses use the full data set. That is, counts of each species at each time point are the cumulative sum of all sites and all years. Records also include net caught specimens and all morphospecies.

### Summary of Results
19 genera had $\ge$ 20 records and these genera are shown for the genus level plot. Additional plots for parasite-host combinations are shown without any record cut-off. Sample size is included in each plot.

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(knitr)
library(ggridges)


# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
dat1 <- read.csv("data/CABS1.csv")


####################################
# Funtions
####################################

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}


```


```{r Data_Wrangling, echo = F, comment = ""}

# general cleaning and conversion:
dat1 <- dat1 %>% 
  uncount(Count) %>% 
  mutate(across(c(Station : Sex, Site), as.factor), # convert to factors
         Date = ymd(Label.Date), # Convert to date
         Year = year(Date), # Create year column
         Month = month(Date), # create month column
         # JD --- Day of year (julian day)
         JD = yday(Date),
         # ToY --- Time of Year
         ToY = case_when(JD < 135 ~ "early",
                         JD > 205 ~ "late",
                         TRUE ~ "mid"),
         # Create a station/year column for grouping
         StationYear = paste(Station, Year, sep = "_")) %>% 
  filter(Station != "",
         Site != "") %>% # remove empty station & site records
  # convert M/F to male/female to avoid "F = FALSE" misinterpretation
  mutate(Sex = case_when(
      Sex == "M" ~ "male",
      Sex == "F" ~ "female",
      TRUE ~ NA_character_ # Handle any unexpected values
    )) %>% 
  # correct misidentification record
  mutate(Short.label.name = case_when(
    Short.label.name == "Megachile apicalis" ~ "Megachile rotundata",
    TRUE ~ as.character(Short.label.name)
  )) %>% 
  # correct misindentification sex classification
  mutate(Sex = case_when(
    Unique.Specimen.Number == 678 |
      Unique.Specimen.Number == 679 ~ "Male",
    TRUE ~ as.character(Sex)
  ))



#############################################

genus.counts <- dat1 %>% 
  select(Genus, Short.label.name) %>% 
  group_by(Genus) %>% 
  distinct() %>% 
  tally(name = "Species")

genus.props <- dat1 %>% 
  group_by(Genus) %>% 
  tally(name = "Prop") %>% 
  mutate(prop.abund = Prop/nrow(dat1))

genus <- left_join(genus.counts, genus.props, by = "Genus")

###################
# color palette
c24 <- c(
  "maroon", "#E31A1C", "green4", "#6A3D9A",
  "#FF7F00", "gold1", "mediumpurple1","skyblue2", "#FB9A99",
  "palegreen2", "#CAB2D6", "#FDBF6F", "gray70", "khaki2",
  "dodgerblue2", "orchid1", "deeppink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3", "darkorange4"
)


genus_levels <- levels(genus$Genus)

color_mapping <- setNames(c24, genus_levels)
```


# Genus by time of year
```{r genus_ridgeplot, echo = F}

genus_records <- dat1 %>% 
  group_by(Genus) %>% 
  tally() %>% filter(n > 19) %>% 
  mutate(stars = case_when(n < 50 & n > 20 ~ "*",
                           n < 100 & n >= 50 ~ "**",
                           n < 1000 & n >= 100 ~ "***",
                           n >= 1000 ~ "****"))

dat_ridges <- dat1 %>% 
  semi_join(genus_records, by = "Genus") %>% 
  mutate(week = week(Date))


genus_orders <- dat_ridges %>% 
  select(Genus, week) %>% 
  group_by(Genus) %>% 
  summarise(peak = Mode(week)) %>% 
  arrange(desc(peak)) %>% 
  data.frame()

level_order <- genus_orders$Genus



dat_ridges %>% 
  ggplot(aes(x = week,
             y = factor(Genus, level = level_order),
             fill = Genus)) +
  geom_density_ridges(aes(height = stat(density)),
                      scale = 5,
                      rel_min_height = 0.01,
                      stat = "density",
                      bw = "bcv") +
  theme_ridges(grid = F) +
  scale_fill_manual(values = color_mapping) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(10, 20, 30, 40),
                     labels = c("3 March", "15 May",
                                "24 July", "3 October")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 12.6,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 25.3,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 38.4,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 7, y = 23, label = " ") +
  annotate(geom = "text", x = 9, y = 21.5, label = "Winter",
           size = 5) +
  annotate(geom = "text", x = 18, y = 21.5, label = "Spring",
           size = 5) +
  annotate(geom = "text", x = 32, y = 21.5, label = "Summer",
           size = 5) +
  annotate(geom = "text", x = 41, y = 21.5, label = "Fall",
           size = 5) +
  geom_text(data = genus_records, aes(x = 43, y = Genus,
                                      label = n), size = 3)


```

**Genus-level seasonal distributions.** Density is estimated at the genus level with Scott's method for genera presumed univoltine, while biased cross validation was used for genera presumed multivoltine. Sample sizes displayed on the right are the total number of records for each genera. Vertical dashed lines correspond to 21 March, 21 June, and 21 September.


### Host-parasite ridges
```{r genus_parasite_host, echo = F}

parasite_records <- dat1 %>% 
  group_by(Genus) %>% 
  tally() %>% 
  filter(Genus == "Nomada" | Genus == "Andrena" |
           Genus == "Agapostemon" | Genus == "Stelis" |
           Genus == "Osmia" | Genus == "Megachile" |
           Genus == "Sphecodes" | Genus == "Halictus" |
           Genus == "Lasioglossum" | Genus == "Epeolus" |
           Genus == "Colletes" | Genus == "Triepeolus" |
           Genus == "Melissodes" | Genus == "Coelioxys")

dat_ridges_parasite <- dat1 %>% 
  semi_join(parasite_records, by = "Genus") %>% 
  mutate(week = week(Date))


parasite_orders <- dat_ridges_parasite %>% 
  select(Genus, JD) %>% 
  group_by(Genus) %>% 
  summarise(peak = Mode(JD)) %>% 
  arrange(desc(peak)) %>% 
  data.frame()

level_order_parasite <- parasite_orders$Genus

```


```{r Nomada_ridges, echo = F}

nomada <- dat_ridges_parasite %>% 
  filter(Genus == "Nomada" | Genus == "Andrena" |
           Genus == "Agapostemon") %>% 
  ggplot(aes(x = week,
             y = factor(Genus, level = c("Agapostemon",
                                         "Andrena",
                                         "Nomada")),
             fill = Genus)) +
  geom_density_ridges(aes(height = stat(density)),
                      scale = 3, 
                      rel_min_height = 0.01,
                      stat = "density",
                      bw = "nrd") +
  theme_ridges(grid = F) +
  scale_fill_manual(values = color_mapping) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(10, 20, 30, 40),
                     labels = c("3 March", "15 May",
                                "24 July", "3 October")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 12.6,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 25.3,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 38.4,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 7, y = 8, label = " ") +
  annotate(geom = "text", x = 9, y = 7, label = "Winter") +
  annotate(geom = "text", x = 18, y = 7, label = "Spring") +
  annotate(geom = "text", x = 32, y = 7, label = "Summer") +
  annotate(geom = "text", x = 41, y = 7, label = "Fall") +
  geom_text(data = filter(parasite_records,
                          Genus == "Nomada" | Genus == "Andrena" |
                          Genus == "Agapostemon"),
            aes(x = 43, y = Genus, label = n), size = 3)

```


```{r Stelis_ridges, echo = F}

stelis <- dat_ridges_parasite %>% 
  filter(Genus == "Stelis" | Genus == "Osmia" |
           Genus == "Megachile") %>% 
  ggplot(aes(x = week,
             y = factor(Genus, level = c("Megachile",
                                         "Osmia",
                                         "Stelis")),
             fill = Genus)) +
  geom_density_ridges(aes(height = stat(density)),
                      scale = 4, 
                      rel_min_height = 0.01,
                      stat = "density",
                      bw = "bcv") +
  theme_ridges(grid = F) +
  scale_fill_manual(values = color_mapping) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(10, 20, 30, 40),
                     labels = c("3 March", "15 May",
                                "24 July", "3 October")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 12.6,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 25.3,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 38.4,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 9, y = 8, label = " ") +
  annotate(geom = "text", x = 10.5, y = 7, label = "Winter") +
  annotate(geom = "text", x = 18, y = 7, label = "Spring") +
  annotate(geom = "text", x = 32, y = 7, label = "Summer") +
  annotate(geom = "text", x = 41, y = 7, label = "Fall") +
  geom_text(data = filter(parasite_records,
                          Genus == "Stelis" | Genus == "Osmia" |
                          Genus == "Megachile"),
            aes(x = 11, y = Genus, label = n), size = 3) +
  scale_y_discrete(position = "right")

##########################################
stelis2 <- dat_ridges_parasite %>% 
  filter(Genus == "Stelis" | Genus == "Osmia" |
           Genus == "Megachile") %>% 
  ggplot(aes(x = week,
             y = factor(Genus, level = c("Megachile",
                                         "Osmia",
                                         "Stelis")),
             fill = Genus)) +
  geom_density_ridges(aes(height = stat(density)),
                      scale = 4, 
                      rel_min_height = 0.01,
                      stat = "density",
                      bw = "bcv") +
  theme_ridges(grid = F) +
  scale_fill_manual(values = color_mapping) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(10, 20, 30, 40),
                     labels = c("3 March", "15 May",
                                "24 July", "3 October")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 12.6,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 25.3,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 38.4,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 9, y = 8, label = " ") +
  annotate(geom = "text", x = 10.5, y = 7, label = "Winter") +
  annotate(geom = "text", x = 18, y = 7, label = "Spring") +
  annotate(geom = "text", x = 32, y = 7, label = "Summer") +
  annotate(geom = "text", x = 41, y = 7, label = "Fall") +
  geom_text(data = filter(parasite_records,
                          Genus == "Stelis" | Genus == "Osmia" |
                          Genus == "Megachile"),
            aes(x = 43, y = Genus, label = n), size = 3)

```


```{r Sphecodes_ridges, echo = F}

sphecodes <- dat_ridges_parasite %>% 
  filter(Genus == "Sphecodes" | Genus == "Halictus" |
           Genus == "Lasioglossum" | Genus == "Agapostemon") %>% 
  ggplot(aes(x = week,
             y = factor(Genus, level = c("Lasioglossum",
                                         "Agapostemon",
                                         "Halictus",
                                         "Sphecodes")),
             fill = Genus)) +
  geom_density_ridges(aes(height = stat(density)),
                      scale = 3, 
                      rel_min_height = 0.01,
                      stat = "density",
                      bw = "bcv") +
  theme_ridges(grid = F) +
  scale_fill_manual(values = color_mapping) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(10, 20, 30, 40),
                     labels = c("3 March", "15 May",
                                "24 July", "3 October")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 12.6,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 25.3,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 38.4,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 6, y = 8, label = " ") +
  annotate(geom = "text", x = 9, y = 7, label = "Winter") +
  annotate(geom = "text", x = 18, y = 7, label = "Spring") +
  annotate(geom = "text", x = 32, y = 7, label = "Summer") +
  annotate(geom = "text", x = 41, y = 7, label = "Fall") +
  geom_text(data = filter(parasite_records,
                          Genus == "Sphecodes" | Genus == "Halictus" |
           Genus == "Lasioglossum" | Genus == "Agapostemon"),
            aes(x = 43, y = Genus, label = n), size = 3)

```


```{r Epeolus_ridges, echo = F}

epeolus <- dat_ridges_parasite %>% 
  filter(Genus == "Epeolus" | Genus == "Colletes") %>% 
  ggplot(aes(x = week,
             y = factor(Genus, level = c("Colletes",
                                         "Epeolus")),
             fill = Genus)) +
  geom_density_ridges(aes(height = stat(density)),
                      scale = 3, 
                      rel_min_height = 0.01,
                      stat = "density",
                      bw = "bcv") +
  theme_ridges(grid = F) +
  scale_fill_manual(values = color_mapping) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(10, 20, 30, 40),
                     labels = c("3 March", "15 May",
                                "24 July", "3 October")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 12.6,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 25.3,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 38.4,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 9, y = 8, label = " ") +
  annotate(geom = "text", x = 10.5, y = 7, label = "Winter") +
  annotate(geom = "text", x = 18, y = 7, label = "Spring") +
  annotate(geom = "text", x = 32, y = 7, label = "Summer") +
  annotate(geom = "text", x = 41, y = 7, label = "Fall") +
  geom_text(data = filter(parasite_records,
                          Genus == "Epeolus" | Genus == "Colletes"),
            aes(x = 11, y = Genus, label = n), size = 3) +
  scale_y_discrete(position = "right")

###############################################3
epeolus2 <- dat_ridges_parasite %>% 
  filter(Genus == "Epeolus" | Genus == "Colletes") %>% 
  ggplot(aes(x = week,
             y = factor(Genus, level = c("Colletes",
                                         "Epeolus")),
             fill = Genus)) +
  geom_density_ridges(aes(height = stat(density)),
                      scale = 3, 
                      rel_min_height = 0.01,
                      stat = "density",
                      bw = "bcv") +
  theme_ridges(grid = F) +
  scale_fill_manual(values = color_mapping) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(10, 20, 30, 40),
                     labels = c("3 March", "15 May",
                                "24 July", "3 October")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 12.6,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 25.3,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 38.4,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 9, y = 8, label = " ") +
  annotate(geom = "text", x = 10.5, y = 7, label = "Winter") +
  annotate(geom = "text", x = 18, y = 7, label = "Spring") +
  annotate(geom = "text", x = 32, y = 7, label = "Summer") +
  annotate(geom = "text", x = 41, y = 7, label = "Fall") +
  geom_text(data = filter(parasite_records,
                          Genus == "Epeolus" | Genus == "Colletes"),
            aes(x = 43, y = Genus, label = n), size = 3)

```


```{r Triepeolus_ridges, echo = F}

triepeolus <- dat_ridges_parasite %>% 
  filter(Genus == "Triepeolus" | Genus == "Melissodes") %>% 
  ggplot(aes(x = week,
             y = factor(Genus, level = c("Melissodes",
                                         "Triepeolus")),
             fill = Genus)) +
  geom_density_ridges(aes(height = stat(density)),
                      scale = 3, 
                      rel_min_height = 0.01,
                      stat = "density",
                      bw = "nrd0") +
  theme_ridges(grid = F) +
  scale_fill_manual(values = color_mapping) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(10, 20, 30, 40),
                     labels = c("3 March", "15 May",
                                "24 July", "3 October")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 12.6,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 25.3,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 38.4,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 7, y = 5, label = " ") +
  annotate(geom = "text", x = 9.5, y = 4.5, label = "Winter") +
  annotate(geom = "text", x = 18, y = 4.5, label = "Spring") +
  annotate(geom = "text", x = 32, y = 4.5, label = "Summer") +
  annotate(geom = "text", x = 45, y = 4.5, label = "Fall") +
  geom_text(data = filter(parasite_records,
                          Genus == "Triepeolus" |
                          Genus == "Melissodes"),
            aes(x = 50, y = c(1, 1.8), label = n), size = 3)



```


```{r Coelioxys_ridges, echo = F}

coelioxys <- dat_ridges_parasite %>% 
  filter(Genus == "Coelioxys" | Genus == "Megachile") %>% 
  ggplot(aes(x = week,
             y = factor(Genus, level = c("Megachile",
                                         "Coelioxys")),
             fill = Genus)) +
  geom_density_ridges(aes(height = stat(density)),
                      scale = 3, 
                      rel_min_height = 0.01,
                      stat = "density",
                      bw = "bcv") +
  theme_ridges(grid = F) +
  scale_fill_manual(values = color_mapping) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(10, 20, 30, 40),
                     labels = c("3 March", "15 May",
                                "24 July", "3 October")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 12.6,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 25.3,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 38.4,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 9, y = 6, label = " ") +
  annotate(geom = "text", x = 10.5, y = 5, label = "Winter") +
  annotate(geom = "text", x = 18, y = 5, label = "Spring") +
  annotate(geom = "text", x = 32, y = 5, label = "Summer") +
  annotate(geom = "text", x = 43, y = 5, label = "Fall") +
  geom_text(data = filter(parasite_records,
                          Genus == "Coelioxys" |
                          Genus == "Megachile"),
            aes(x = 11, y = Genus, label = n), size = 3) +
  scale_y_discrete(position = "right")

coelioxys2 <- dat_ridges_parasite %>% 
  filter(Genus == "Coelioxys" | Genus == "Megachile") %>% 
  ggplot(aes(x = week,
             y = factor(Genus, level = c("Megachile",
                                         "Coelioxys")),
             fill = Genus)) +
  geom_density_ridges(aes(height = stat(density)),
                      scale = 3, 
                      rel_min_height = 0.01,
                      stat = "density",
                      bw = "bcv") +
  theme_ridges(grid = F) +
  scale_fill_manual(values = color_mapping) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(10, 20, 30, 40),
                     labels = c("3 March", "15 May",
                                "24 July", "3 October")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 12.6,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 25.3,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 38.4,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 6, y = 6, label = " ") +
  annotate(geom = "text", x = 9, y = 5, label = "Winter") +
  annotate(geom = "text", x = 18, y = 5, label = "Spring") +
  annotate(geom = "text", x = 32, y = 5, label = "Summer") +
  annotate(geom = "text", x = 43, y = 5, label = "Fall") +
  geom_text(data = filter(parasite_records,
                          Genus == "Coelioxys" |
                          Genus == "Megachile"),
            aes(x = 45, y = Genus, label = n), size = 3)

```

### Parasite-host composite
```{r parasite_host_composite, echo = F, fig.width = 12, fig.height = 10}

plot_grid(nomada, stelis, sphecodes,
          epeolus, triepeolus, coelioxys,
          labels = c("A", "B", "C", "D", "E", "F"),
          nrow = 3, rel_widths = c(0.9, 1),
          greedy = FALSE)


```

**Genus-level seasonal distributions** for the parasites: A) Nomada, B) Stelis, C) Sphecodes, D) Epeolus, E) Triepeolus, and F) Coelioxys. Beneath each parasite genera are the presumed host genera. Density is estimated at the genus level with Scott's method for genera presumed univoltine, while biased cross validation was used for genera presumed multivoltine. Sample sizes displayed on the right (plots A, C, E) and left (plots B, D, F) are the total number of records for each genera. Vertical dashed lines correspond to 21 March, 21 June, and 21 September.


### Parasite-host composite (3 good combos)
```{r para_host_composite_3, echo = F, fig.width = 8, fig.height = 7}

plot_grid(nomada, sphecodes, coelioxys2,
          labels = c("A", "B", "C"),
          nrow = 3,
          greedy = FALSE)


```

**Genus-level seasonal distributions** for the parasites: A) Nomada, B) Sphecodes, and C) Coelioxys. Beneath each parasite genera are the presumed host genera. Density is estimated at the genus level with Scott's method for genera presumed univoltine, while biased cross validation was used for genera presumed multivoltine. Sample sizes displayed on the right are the total number of records for each genera. Vertical dashed lines correspond to 21 March, 21 June, and 21 September.


### Parasite-host composite (3 bad combos)
```{r parasite_host_composite_3_bad, echo = F, fig.width = 8, fig.height = 7}

plot_grid(stelis2, epeolus2, triepeolus,
          labels = c("A", "B", "C"),
          nrow = 3,
          greedy = FALSE)


```

**Genus-level seasonal distributions** for the parasites: A) Stelis, B) Epeolus, and C) Triepeolus. Beneath each parasite genera are the presumed host genera. Density is estimated at the genus level with Scott's method for genera presumed univoltine, while biased cross validation was used for genera presumed multivoltine. Sample sizes displayed on the right are the total number of records for each genera. Vertical dashed lines correspond to 21 March, 21 June, and 21 September. These parasites have low sample sizes limiting distributional estimation.



# Species by time of year
```{r species_ridgeplot, echo = F, fig.height = 10}

more_than_20 <- dat1 %>% group_by(Short.label.name) %>% 
  tally() %>% 
  filter(n > 20)

dat_ridges_species <- dat1 %>%
  mutate(week = week(Date)) %>% 
  semi_join(more_than_20, by = "Short.label.name")



species_orders <- dat_ridges_species %>% 
  select(Short.label.name, JD) %>% 
  group_by(Short.label.name) %>% 
  summarise(peak = Mode(JD)) %>% 
  arrange(desc(peak)) %>% 
  data.frame()

level_order_species <- species_orders$Short.label.name


dat_ridges_species %>% 
  ggplot(aes(x = week,
             y = factor(Short.label.name,
                        level = level_order_species),
             fill = Short.label.name)) +
  geom_density_ridges(scale = 3, rel_min_height = 0.01) +
  theme_ridges(grid = F) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(10, 20, 30, 40),
                     labels = c("3 March", "15 May",
                                "24 July", "3 October")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 12.6,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 25.3,
             color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 38.4,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 7, y = 54, label = " ") +
  annotate(geom = "text", x = 8, y = 52, label = "Winter") +
  annotate(geom = "text", x = 20, y = 52, label = "Spring") +
  annotate(geom = "text", x = 32, y = 52, label = "Summer") +
  annotate(geom = "text", x = 42, y = 52, label = "Fall") +
  geom_text(data = more_than_20, aes(x = 43, y = Short.label.name,
                                     label = n), size = 3)


```

**Species-level seasonal distributions.** Density is estimated uniformly across all species with Silverman's method. Sample sizes displayed on the right are the total number of records for each species. Species displayed are those for which sample sizes were $\ge$ 20. Vertical dashed lines correspond to 21 March, 21 June, and 21 September.

## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


