---
title: "BeeSearch Sex Ratio Analyses"
author: "Dr. Riley M. Anderson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = F,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

This analysis looks at sex ratio bias in sweep net traps vs. blue vane traps.

### Summary of Results


```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(vegan)
library(adespatial)
library(knitr)
library(fossil)
library(geosphere)
library(randomForest)
library(caret)



# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
dat1 <- read.csv("data/CABS1.csv")

comp.sites <- read.csv("data/comparable.sites.csv")

site.locs <- read.csv("data/site.locations.csv")

temps <- read.csv("data/temperatures.csv")
####################################
# Functions
####################################

sex_ratio_function <- function(data, group){
  data  %>%
    group_by(across(all_of(group))) %>%
    tally(name = "count") %>%
    pivot_wider(names_from = Sex,
                values_from = count,
                values_fill = list(count = 0)) %>%
    mutate(sex_ratio = male/female)
}

###################################
###################
# color palette
c24 <- c(
  "maroon", "#E31A1C", "green4", "#6A3D9A",
  "#FF7F00", "gold1", "mediumpurple1","skyblue2", "#FB9A99",
  "palegreen2", "#CAB2D6", "#FDBF6F", "gray70", "khaki2",
  "dodgerblue2", "orchid1", "deeppink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3", "darkorange4"
)


```


```{r Data_Wrangling, echo = F, comment = ""}

# general cleaning and conversion:
dat1 <- dat1 %>% 
  uncount(Count) %>% 
  mutate(across(c(Station : Sex, Site), as.factor), # convert to factors
         Date = ymd(Label.Date), # Convert to date
         Year = year(Date), # Create year column
         Month = month(Date), # create month column
         # JD --- Day of year (julian day)
         JD = yday(Date),
         # ToY --- Time of Year
         ToY = case_when(JD < 135 ~ "early",
                         JD > 205 ~ "late",
                         TRUE ~ "mid"),
         # Create a station/year column for grouping
         StationYear = paste(Station, Year, sep = "_")) %>% 
  filter(Station != "",
         Site != "") %>% # remove empty station & site records
  # convert M/F to male/female to avoid "F = FALSE" misinterpretation
  mutate(Sex = case_when(
      Sex == "M" ~ "male",
      Sex == "F" ~ "female",
      TRUE ~ NA_character_ # Handle any unexpected values
    )) %>% 
  # correct misidentification record
  mutate(Short.label.name = case_when(
    Short.label.name == "Megachile apicalis" ~ "Megachile rotundata",
    TRUE ~ as.character(Short.label.name)
  )) %>% 
  # correct misindentification sex classification
  mutate(Sex = case_when(
    Unique.Specimen.Number == 678 |
      Unique.Specimen.Number == 679 ~ "Male",
    TRUE ~ as.character(Sex)
  ))


# Create data with all the incompatible sites (unequal sampling effort)
incomp.sites <- dat1 %>% 
  select(StationYear) %>% 
  filter(StationYear == "POS1_2018" |
           StationYear == "POS10_2019" |
           StationYear == "POS11_2018" |
           StationYear == "POS13_2017" |
           StationYear == "POS14_2017" |
           StationYear == "POS15_2017" |
           StationYear == "POS16_2017" |
           StationYear == "POS17_2017" |
           StationYear == "POS18_2017" |
           StationYear == "POS19_2017" |
           StationYear == "POS20_2017" |
           StationYear == "POS20_2019" |
           StationYear == "POS21_2017" |
           StationYear == "POS22_2019" |
           StationYear == "POS23_2019" |
           StationYear == "POS1_2014" |
           StationYear == "POS2_2014" |
           StationYear == "POS3_2014") %>% 
  distinct() %>% 
  mutate(StationYear = factor(StationYear))


# filter out the sites with incompatible sampling effort
dat2 <- anti_join(dat1, incomp.sites, by = "StationYear") %>% 
  # unequal sampling in 2020
  filter(Year != 2020) %>% 
  mutate(Year = factor(Year)) %>% 
  droplevels()

# keep the single record of morphospecies "Hylaeus sp. 2023_1"
keep.Hylaeus <- dat2 %>% 
  filter(str_detect(Short.label.name, "Hylaeus sp. 2023"))

# remove any records where species is undetermined:
dat2 <- dat2 %>% 
  filter(!str_detect(Short.label.name, "(?<!m)sp\\.")) %>% 
  bind_rows(keep.Hylaeus)
# ^^^ this data will be used when analyses do not depend on number of species


# Filter out certain morphospecies that could inflate richness
dat.ex.msp <- dat2 %>% 
  filter(!(Sex == "male" & str_detect(Short.label.name,
                                    "Sphecodes msp\\.")),
         !(Sex == "female" & str_detect(Short.label.name,
                                       "Osmia msp\\.")),
         !(Sex == "female" & str_detect(Short.label.name,
                                        "Nomada msp\\.")))
# ^^^ this data will be used for analyses involving species counts


######################################################
```

# Sex ratio in trap vs. netting collection method:

* Ratios > 1 represent male bias
* Ratios < 1 represent female bias


* Overall:
```{r overall_sex_ratios, echo = F}
kable(sex_ratio_function(dat2, c("Sex", "Collection.Method")),
      format = "markdown", digits = 2)
```

* By site:
```{r sex_ratios_by_site, echo = F}
kable(sex_ratio_function(dat2, c("Sex", "Collection.Method", "Site")),
      format = "markdown", digits = 2)

sex_ratio_function(dat2, c("Sex", "Collection.Method", "Site")) %>% 
  ggplot(aes(x = Site, y = sex_ratio, fill = Collection.Method)) +
  geom_col(position = position_dodge()) +
  labs(y = "Sex ratio",
       fill = "Collection\nmethod") +
  scale_fill_discrete(labels = c("Net", "Trap")) +
  theme_cowplot() +
  geom_hline(yintercept = 1,
             linetype = "dashed")

```
**Sex ratios of bees by site and collection method.** Points above the dashed line represent male bias, whereas point below represent female bias.


* By season:
```{r sex_ratios_by_season, echo = F}
kable(sex_ratio_function(dat2, c("Sex", "Collection.Method", "ToY")),
      format = "markdown", digits = 2)
```

* By season and site:
```{r sex_ratios_by_season_and_site, echo = F}
kable(sex_ratio_function(dat2, c("Sex", "Collection.Method",
                                 "ToY", "Site")),
      format = "markdown", digits = 2)

sex_ratio_function(dat2, c("Sex",
                           "Collection.Method",
                           "ToY")) %>% 
  ggplot(aes(x = ToY, y = sex_ratio)) +
  geom_point(aes(fill = Collection.Method),
             size = 3,
             shape = 21) +
  scale_x_discrete(limits = c("early", "mid", "late")) +
  geom_jitter(data = filter(sex_ratio_function(dat2, c("Sex",
                                               "Collection.Method",
                                               "ToY", "Site")),
                            sex_ratio < 5),
             aes(x = ToY, y = sex_ratio, color = Collection.Method),
             width = 0.2,
             show.legend = F,
             shape = 21) +
  labs(x = "Time of season",
       y = "Sex ratio",
       fill = "Collection\nmethod") +
  scale_fill_discrete(labels = c("Net", "Trap")) +
  scale_y_continuous(limits = c(0, 5)) +
  theme_cowplot() +
  geom_hline(yintercept = 1,
             linetype = "dashed")

```
**Sex ratios of bees collected by either net or trap across the sampling season.** Points above the dashed line represent male bias, whereas point below represent female bias. Early season (late March - mid May), mid season (mid May to mid July), and late season (mid July - late September) time windows in the sampling effort are shown. Large solid points are sex ratios calculated at the season level. Small open points are sex ratios calculated at the sub-site level within each season. Small points are spread horizontally for visual clarity.


## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


